---
title: "`r params$title`"
author: "`r params$author`"
date: "`r Sys.Date()`"
params:
    callset: "MISSING"
    author: "Davis McCarthy"
    title: "Differential gene and pathway analysis between clones"
    to_working_dir: "./"

output: 
    html_document:
        toc: true
        toc_float: true
        theme: journal
        highlight: pygments
        number_sections: true
        code_folding: hide

---

# Differential gene and pathway expression analysis between clones: unstimulated cells

**Cardelino cell assignments using Canopy trees**. No merging of clones.

Using the latest callset from Petr Danecek (April 2018): fisher exact test and 
other filters. This analysis uses the callset: `r params$callset`.

```{r global_options, include=FALSE}
library(vcfR)
library(readr)
library(dplyr)
library(scran)
library(scater)
library(viridis)
library(ggplot2)
library(ggridges)
library(SingleCellExperiment)
library(edgeR)
library(limma)
library(org.Hs.eg.db)
library(cowplot)
library(gplots)
knitr::opts_chunk$set(fig.width=10, fig.height=7,
                      echo=TRUE, warning=FALSE, message=TRUE)
options(stringsAsFactors = FALSE)
```

## Load data

Define the donors for which we have cell assignment results to study.

```{r define-donors}
cat(getwd())
fls <- list.files(file.path(to_working_dir, "Data/processed/sces"))
fls <- fls[grepl(params$callset, fls)]
donors <- gsub(".*ce_([a-z]+)_.*", "\\1", fls)
length(donors)
```

Read in annotated SCE objects and create a list of SCE objects containing all 
cells (stimulated and unstimulated) and just unstimulated cells.

```{r load-sces}
sce_unst_list <- list()
for (don in donors) {
    sce_unst_list[[don]] <- readRDS(file.path(to_working_dir, "Data/processed/sces", 
        paste0("sce_", don, "_with_clone_assignments.", params$callset, ".rds")))
    cat(paste("reading", don, ":   ", ncol(sce_unst_list[[don]]), "unstimulated cells.\n"))
}
```

We have `r length(sce_unst_list)` donors for analysis.


## Generate SCE list with only assigned cells in clones with more than three cells

Filter donors for DE/pathway analysis using all (stimulated and unstimulated cells).

* Keep clones with at least three assigned cells.
* Keep only cells that are confidently assigned to a (kept) clone. 
* Keep donors that have at least 15 assigned cells to kept clones

Conduct the filtering for the unstimulated cells only.

```{r filter-donors}
sce_unst_list_filt <- list()
clones_keep_unst <- list()
for (i in names(sce_unst_list) ) {
    clones_keep_unst[[i]] <- table(sce_unst_list[[i]]$assigned)[which(table(sce_unst_list[[i]]$assigned) >= 3)]
    clones_keep_unst[[i]] <- clones_keep_unst[[i]][!(names(clones_keep_unst[[i]]) %in% "unassigned")]
    cells_use_unst <- colnames(sce_unst_list[[i]])[sce_unst_list[[i]]$assigned %in% names(clones_keep_unst[[i]])]
    cat(paste("....", i, ":", length(cells_use_unst), "unstimulated cells for DE analyses.\n"))
    if (length(cells_use_unst) > 14.5 && length(clones_keep_unst[[i]]) > 1.5)
        sce_unst_list_filt[[i]] <- sce_unst_list[[i]][, cells_use_unst]
}
```

```{r}
length(sce_unst_list_filt)
```

`r length(sce_unst_list_filt)` donors are retained for DE/pathway analysis using
unstimulated cells only.

Plot assignment summaries for filtered donors.

Define data frames with numbers of cells assigned for different clones by donor.

```{r}
cells_per_clone_unst_filt <- list()
for (i in names(sce_unst_list_filt)) {
    if (ncol(sce_unst_list_filt[[i]]) == 0)
        cat(paste0("No unstimulated cells for ", i, "\n"))
    else {
        tmp <- as_data_frame(table(sce_unst_list_filt[[i]]$assigned))
        colnames(tmp) <- c("clone", "n_cells")
        tmp[["donor"]] <- i
        tmp2 <- data.frame(clone = "total assigned", 
                           n_cells = sum(tmp[["n_cells"]][tmp[["clone"]] != "unassigned"]), donor = i)
        tmp <- bind_rows(tmp, tmp2)
        tmp[["assigned"]] <- tmp[["clone"]] != "unassigned"
        cells_per_clone_unst_filt[[i]] <- tmp
    }
}
df_cells_per_clone_unst_filt <- do.call("bind_rows", cells_per_clone_unst_filt)

coldata_list <- list()
coldata_filt_list <- list()
for (i in names(sce_unst_list_filt)) {
    coldata_list[[i]] <- as_data_frame(
                    as.data.frame(colData(sce_unst_list[[i]])))
    coldata_list[[i]]$donor <- i
    coldata_filt_list[[i]] <- as_data_frame(
                    as.data.frame(colData(sce_unst_list_filt[[i]])))
    coldata_filt_list[[i]]$donor <- i
    
}
df_coldata <- do.call("bind_rows", coldata_list)
df_coldata_filt <- do.call("bind_rows", coldata_filt_list)
```

`r nrow(df_coldata)` QC-passing cells were obtained for the donors used in this
DE analysis. `r 100 * mean(df_coldata$assigned != "unassigned")`% of these cells were 
assigned to a donor. After filtering out unassigned cells and clones with very
small cell numbers, `r nrow(df_coldata_filt)` cells remain for analysis.

Assignment rates by donor:

```{r assignment-rates}
dplyr::group_by(df_coldata, donor) %>% 
        dplyr::summarise(prop_assigned = mean(assigned != "unassigned")) %>% 
        print(n=Inf)
```


```{r}
options(repr.plot.width = 9, repr.plot.height = 9)
ggplot(df_cells_per_clone_unst_filt, aes(x = n_cells, y = reorder(donor, n_cells, max), 
        colour = clone, shape = assigned)) +
geom_point(size = 4, alpha = 0.7) +
scale_shape_manual(values = c(4, 19)) +
scale_color_manual(values = c(magma(8)[-c(1,8)], "gray30", "black")) +
#ggthemes::scale_colour_tableau() +
theme_bw(16) + xlab("Number of cells") + ylab("Donor") +
ggtitle("Cells per clone, by donor")

```


## DE analysis for each donor - unstimulated cells

```{r}
data_dir <- file.path(to_working_dir, "Data/processed/de_analysis_FTv62/cellcycle_analyses/")
fig_dir <- file.path(to_working_dir, "figures/clonality/de_pathways/cellcycle_analyses/")
```

```{r filter-genes}
gene_use_DE <- (!rowData(sce_unst_list_filt[[1]])$is_feature_control) # remove ERCC
for (i in names(sce_unst_list_filt)) {
  gene_use_DE <- gene_use_DE & (rowMeans(counts(sce_unst_list_filt[[i]]) > 1)) # all 0s
  # 1 count in 10% of cells
  # rowMeans of counts > 0.5
}
```

```{r}
sum(gene_use_DE)
```

```{r fit-qlf-models}
dge_list_unst <- list()
dge_list_unst_cellcycle_only <- list()
dge_list_unst_cellcycle_clone <- list()
design_list_unst_cellcycle_only <- list()
design_list_unst_cellcycle_clone <- list()
fit_list_unst_cellcycle_only <- list()
fit_list_unst_cellcycle_clone <- list()
qlf_list_unst_cellcycle_only <- list()
qlf_list_unst_cellcycle_clone <- list()

for (i in names(sce_unst_list_filt)) {
  #sce_list_unst[[i]] <- scran::computeSumFactors(sce_list_unst[[i]])
    cat("....calculating DE for ", i, "\n")
    dge_list_unst[[i]] <- edgeR::DGEList(round(counts(sce_unst_list_filt[[i]][gene_use_DE,])))
    dge_list_unst[[i]] <- edgeR::calcNormFactors(dge_list_unst[[i]], method = "TMM")
    sce_unst_list_filt[[i]]$cdr <- (colSums(counts(sce_unst_list_filt[[i]]) > 0) / nrow(sce_unst_list_filt[[i]]))
    sce_unst_list_filt[[i]]$plate <- as.factor(sce_unst_list_filt[[i]]$plate)
    design_list_unst_cellcycle_only[[i]] <- model.matrix(~cdr + plate + G1 + S + G2M, 
                                        data = colData(sce_unst_list_filt[[i]]))
    design_list_unst_cellcycle_clone[[i]] <- model.matrix(~cdr + plate + G1 + S + G2M + assigned, 
                                        data = colData(sce_unst_list_filt[[i]]))

}

## cell cycle only
for (i in names(sce_unst_list_filt)) {
    dge_list_unst_cellcycle_only[[i]] <- estimateDisp(
        dge_list_unst[[i]], design_list_unst_cellcycle_only[[i]])
    fit_list_unst_cellcycle_only[[i]] <- glmQLFit(
        dge_list_unst_cellcycle_only[[i]], design_list_unst_cellcycle_only[[i]])
    ncoef <- ncol(design_list_unst_cellcycle_only[[i]])
    qlf_list_unst_cellcycle_only[[i]] <- glmQLFTest(
        fit_list_unst_cellcycle_only[[i]], 
        coef = (ncoef - 2):ncoef)
    sum(p.adjust(qlf_list_unst_cellcycle_only[[i]]$table$PValue, 
                    method = "BH") <= 0.05, na.rm = TRUE)
    summary(decideTestsDGE(qlf_list_unst_cellcycle_only[[i]]))
}

## test clone DE accounting for cell cycle
for (i in names(sce_unst_list_filt)) {
    dge_list_unst_cellcycle_clone[[i]] <- estimateDisp(
        dge_list_unst[[i]], design_list_unst_cellcycle_clone[[i]])
    fit_list_unst_cellcycle_clone[[i]] <- glmQLFit(
        dge_list_unst_cellcycle_clone[[i]], design_list_unst_cellcycle_clone[[i]])
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    qlf_list_unst_cellcycle_clone[[i]] <- glmQLFTest(
        fit_list_unst_cellcycle_clone[[i]],
        coef = (ncol(design_list_unst_cellcycle_clone[[i]]) - num_clones + 2):ncol(design_list_unst_cellcycle_clone[[i]]))
    sum(p.adjust(qlf_list_unst_cellcycle_clone[[i]]$table$PValue, 
                    method = "BH") <= 0.05, na.rm = TRUE)
    summary(decideTestsDGE(qlf_list_unst_cellcycle_clone[[i]]))
}


rm(dge_list_unst)
```


### Conduct all pairwise tests for each donor

Conduct QL F-test for each possible pairwise contrast between clones when 
accounting for cell cycle scores.

```{r de-pairwise-contrasts}
qlf_pairwise_list <- list()
for (i in names(dge_list_unst_cellcycle_clone)) {
    cat("....calculating DE for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    num_covars <- ncol(design_list_unst_cellcycle_clone[[i]]) - (num_clones - 1)
    num_contrs <- choose(num_clones, 2)
    out_list <- list()
    out_list$base_clone <- levels(factor(sce_unst_list_filt[[i]]$assigned))[1]
    for (j in 2:num_clones) {
        coef_idx <- (ncol(design_list_unst_cellcycle_clone[[i]]) - num_clones + j)
        tmp <- glmQLFTest(fit_list_unst_cellcycle_clone[[i]], coef = coef_idx)
        comp <- paste0(
            gsub("assigned", "",
                 colnames(design_list_unst_cellcycle_clone[[i]])[coef_idx]),  
            "_",
            out_list$base_clone)
        out_list[[comp]] <- tmp
    }
    if (num_clones > 2) {
        contrasts <- list()
        n <- 1
        for (k in seq_len(num_clones - 1)) {
            for (m in seq_len(num_clones - 1)) {
                if (m < k) {
                    contr <- rep(0, ncol(design_list_unst_cellcycle_clone[[i]]))
                    contr[num_covars + m] <- -1
                    contr[num_covars + k] <- 1
                    contrasts[[n]] <- contr
                    n <- n + 1
                }
            }
        }
        for (p in seq_len(length(contrasts))) {
            tmp <- glmQLFTest(fit_list_unst_cellcycle_clone[[i]], contrast = contrasts[[p]])
            tmp2 <- strsplit(tmp$comparison, split = " ")[[1]]
            comp <- paste0(
                gsub("1\\*assigned", "", tmp2[2]),
                "_",
                gsub("-1\\*assigned", "", tmp2[1]))
            out_list[[comp]] <- tmp
        }
    }
    qlf_pairwise_list[[i]] <- out_list
}
```


```{r}
pryr::mem_used()
```



## Gene set testing

Look for enriched gene sets for clone vs base clone logFCs accounting for cell
cycle scores from cyclone.

```{r}
load(file.path(to_working_dir, "Data/human_c6_v5p2.rdata"))
load(file.path(to_working_dir, "Data/human_H_v5p2.rdata"))
load(file.path(to_working_dir, "Data/human_c2_v5p2.rdata"))
```

```{r}
#For the reverse map ENSEMBL2EG:
# Convert to a list
xx <- as.list(org.Hs.egENSEMBL2EG)
```

### Oncogenic gene sets (c6)

Testing enrichment of oncogenic gene sets (c6) for all pairwise contrasts.

```{r genesets-c6}
camera_msigdb_c6_list <- list()
for (i in names(sce_unst_list_filt)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    camera_msigdb_c6_list[[i]] <- list()
    for (j in 2:length(qlf_pairwise_list[[i]])) {
        qlf_pairwise_list[[i]][[j]]$table$ensembl_gene_id <- 
            strsplit2(rownames(qlf_pairwise_list[[i]][[j]]$table), split = "_")[,1]
        qlf_pairwise_list[[i]][[j]]$table$hgnc_symbol <- 
            strsplit2(rownames(qlf_pairwise_list[[i]][[j]]$table), split = "_")[,2]
        qlf_pairwise_list[[i]][[j]]$table$entrezid <- NA
        for (k in seq_len(nrow(qlf_pairwise_list[[i]][[j]]$table))) {
            if (qlf_pairwise_list[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                qlf_pairwise_list[[i]][[j]]$table$entrezid[k] <- xx[[qlf_pairwise_list[[i]][[j]]$table$ensembl_gene_id[k]]][1]
        }
        idx <- ids2indices(Hs.c6, 
                           identifiers = qlf_pairwise_list[[i]][[j]]$table$entrezid)
        camera_msigdb_c6_list[[i]][[names(qlf_pairwise_list[[i]])[j]]] <- list()
        camera_msigdb_c6_list[[i]][[names(qlf_pairwise_list[[i]])[j]]][["logFC"]] <- 
            cameraPR(statistic = qlf_pairwise_list[[i]][[j]]$table$logFC, idx, 
                     inter.gene.cor = 0.01)
        camera_msigdb_c6_list[[i]][[names(qlf_pairwise_list[[i]])[j]]][["signF"]] <- 
            cameraPR(statistic = (sign(qlf_pairwise_list[[i]][[j]]$table$logFC) * 
                                      qlf_pairwise_list[[i]][[j]]$table$F), idx, 
                     inter.gene.cor = 0.01)
    }
}
```


### Curated gene sets (c2)

Testing enrichment of curated gene sets (c2) for all pairwise contrasts.

```{r genesets-c2}
camera_msigdb_c2_list <- list()
for (i in names(sce_unst_list_filt)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    camera_msigdb_c2_list[[i]] <- list()
    for (j in  2:length(qlf_pairwise_list[[i]])) {
        qlf_pairwise_list[[i]][[j]]$table$ensembl_gene_id <- 
            strsplit2(rownames(qlf_pairwise_list[[i]][[j]]$table), split = "_")[,1]
        qlf_pairwise_list[[i]][[j]]$table$hgnc_symbol <- 
            strsplit2(rownames(qlf_pairwise_list[[i]][[j]]$table), split = "_")[,2]
        qlf_pairwise_list[[i]][[j]]$table$entrezid <- NA
        for (k in seq_len(nrow(qlf_pairwise_list[[i]][[j]]$table))) {
            if (qlf_pairwise_list[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                qlf_pairwise_list[[i]][[j]]$table$entrezid[k] <- xx[[qlf_pairwise_list[[i]][[j]]$table$ensembl_gene_id[k]]][1]
        }
        idx <- ids2indices(Hs.c2, 
                           identifiers = qlf_pairwise_list[[i]][[j]]$table$entrezid)
        camera_msigdb_c2_list[[i]][[names(qlf_pairwise_list[[i]])[j]]] <- list()
        camera_msigdb_c2_list[[i]][[names(qlf_pairwise_list[[i]])[j]]][["logFC"]] <- 
            cameraPR(statistic = qlf_pairwise_list[[i]][[j]]$table$logFC, idx, 
                     inter.gene.cor = 0.01)
        camera_msigdb_c2_list[[i]][[names(qlf_pairwise_list[[i]])[j]]][["signF"]] <- 
            cameraPR(statistic = (sign(qlf_pairwise_list[[i]][[j]]$table$logFC) * 
                                      qlf_pairwise_list[[i]][[j]]$table$F), idx, 
                     inter.gene.cor = 0.01)
    }
}
```



### Hallmark gene sets (H)

Testing enrichment of Hallmark gene sets (H) for all pairwise contrasts.

```{r genesets-H}
camera_msigdb_H_list <- list()
for (i in names(sce_unst_list_filt)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    camera_msigdb_H_list[[i]] <- list()
    for (j in 2:length(qlf_pairwise_list[[i]])) {
        qlf_pairwise_list[[i]][[j]]$table$ensembl_gene_id <- 
            strsplit2(rownames(qlf_pairwise_list[[i]][[j]]$table), split = "_")[,1]
        qlf_pairwise_list[[i]][[j]]$table$hgnc_symbol <- 
            strsplit2(rownames(qlf_pairwise_list[[i]][[j]]$table), split = "_")[,2]
        qlf_pairwise_list[[i]][[j]]$table$entrezid <- NA
        for (k in seq_len(nrow(qlf_pairwise_list[[i]][[j]]$table))) {
            if (qlf_pairwise_list[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                qlf_pairwise_list[[i]][[j]]$table$entrezid[k] <- xx[[qlf_pairwise_list[[i]][[j]]$table$ensembl_gene_id[k]]][1]
        }
        idx <- ids2indices(Hs.H, 
                           identifiers = qlf_pairwise_list[[i]][[j]]$table$entrezid)
        camera_msigdb_H_list[[i]][[names(qlf_pairwise_list[[i]])[j]]] <- list()
        camera_msigdb_H_list[[i]][[names(qlf_pairwise_list[[i]])[j]]][["logFC"]] <- 
            cameraPR(statistic = qlf_pairwise_list[[i]][[j]]$table$logFC, idx, 
                     inter.gene.cor = 0.01)
        camera_msigdb_H_list[[i]][[names(qlf_pairwise_list[[i]])[j]]][["signF"]] <- 
            cameraPR(statistic = (sign(qlf_pairwise_list[[i]][[j]]$table$logFC) * 
                                      qlf_pairwise_list[[i]][[j]]$table$F), idx, 
                     inter.gene.cor = 0.01)
    }
}
```


## Save results to file

Initialise results list.

```{r}
de_results_list_unst <- list()
```

Add *camera* results.

```{r}
de_results_list_unst[["camera"]] <- list()
de_results_list_unst[["camera"]][["c2"]] <- camera_msigdb_c2_list
de_results_list_unst[["camera"]][["c6"]] <- camera_msigdb_c6_list
de_results_list_unst[["camera"]][["H"]] <- camera_msigdb_H_list
```

Add edgeR QL F-test results.

```{r}
de_results_list_unst[["cellcycle_only"]] <- list()
de_results_list_unst[["cellcycle_only"]][["design_list"]] <- design_list_unst_cellcycle_only
de_results_list_unst[["cellcycle_only"]][["dge_list"]] <- dge_list_unst_cellcycle_only
de_results_list_unst[["cellcycle_only"]][["fit_list"]] <- fit_list_unst_cellcycle_only
de_results_list_unst[["cellcycle_only"]][["qlf_list"]] <- qlf_list_unst_cellcycle_only
de_results_list_unst[["cellcycle_clone"]] <- list()
de_results_list_unst[["cellcycle_clone"]][["design_list"]] <- design_list_unst_cellcycle_clone
de_results_list_unst[["cellcycle_clone"]][["dge_list"]] <- dge_list_unst_cellcycle_clone
de_results_list_unst[["cellcycle_clone"]][["fit_list"]] <- fit_list_unst_cellcycle_clone
de_results_list_unst[["cellcycle_clone"]][["qlf_list"]] <- qlf_list_unst_cellcycle_clone
de_results_list_unst[["cellcycle_clone"]][["qlf_pairwise"]] <- qlf_pairwise_list
de_results_list_unst[["sce_list_unst"]] <- sce_unst_list_filt
```

```{r}
data_dir
```

```{r}
saveRDS(object = de_results_list_unst, 
    file = file.path(data_dir, 
    paste0(params$callset, ".de_results_unstimulated_cells.rds")))
```

```{r}
pryr::object_size(de_results_list_unst)
```

## BCV and MD plots

```{r}
for (i in names(sce_unst_list_filt)) {
    cat("....plotting BCV/MD plots for ", i, "\n")
    png(file.path(fig_dir, paste0("BCV_plots/", params$callset, ".cellcycle_clone.CV_", i, ".png")))
    plotBCV(dge_list_unst_cellcycle_clone[[i]])
    dev.off()
    png(file.path(fig_dir, paste0("BCV_plots/", params$callset, ".cellcycle_only.CV_", i, ".png")))
    plotBCV(dge_list_unst_cellcycle_only[[i]])
    dev.off()
    png(file.path(fig_dir, paste0("MD_plots/", params$callset, ".cellcycle_clone.MD_", i, ".png")))
    plotMD(qlf_list_unst_cellcycle_clone[[i]], main = paste(i, " DE"))
    dev.off()
    png(file.path(fig_dir, paste0("MD_plots/", params$callset, ".cellcycle_only.MD_", i, ".png")))
    plotMD(qlf_list_unst_cellcycle_only[[i]], main = paste(i, " DE"))
    dev.off()
}
```



# Session info

```{r sessioninfo}
devtools::session_info()
```

