---
title: "`r params$title`"
author: "`r params$author`"
date: "`r Sys.Date()`"
params:
    callset: "MISSING"
    author: "Davis McCarthy"
    title: "Differential gene and pathway analysis between clones"
    to_working_dir: "./"

output: 
    html_document:
        toc: true
        toc_float: true
        theme: journal
        highlight: pygments
        number_sections: true
        code_folding: hide

---

# Differential gene and pathway expression analysis between clones: unstimulated cells

**Permuting clones** to get empirical p-values for DE and pathway effects.

**Cardelino cell assignments using Canopy trees**. No merging of clones.

Using the latest callset from Petr Danecek (April 2018): fisher exact test and 
other filters. This analysis uses the callset: `r params$callset`.

```{r global_options, include=FALSE}
library(vcfR)
library(readr)
library(dplyr)
library(scran)
library(scater)
library(viridis)
library(ggplot2)
library(ggridges)
library(SingleCellExperiment)
library(edgeR)
library(limma)
library(org.Hs.eg.db)
library(cowplot)
library(gplots)
library(goseq)
```

## Load data

Define the donors for which we have cell assignment results to study.

```{r}
set.seed(20180520)
nperm <- 100

cat(getwd())
fls <- list.files(file.path(to_working_dir, "Data/processed/sces"))
fls <- fls[grepl(params$callset, fls)]
donors <- gsub(".*ce_([a-z]+)_.*", "\\1", fls)
length(donors)
```

Read in annotated SCE objects and create a list of SCE objects containing all 
cells (stimulated and unstimulated) and just unstimulated cells.

```{r}
sce_list <- list()
sce_unst_list <- list()
for (don in donors) {
    sce_list[[don]] <- readRDS(file.path(to_working_dir, "Data/processed/sces", 
        paste0("sce_", don, "_with_clone_assignments.", params$callset, ".rds")))
    sce_unst_list[[don]] <- sce_list[[don]][, sce_list[[don]]$well_condition == "unstimulated"]
    cat(paste("reading", don, ":   ", ncol(sce_list[[don]]), "cells total,", 
        ncol(sce_unst_list[[don]]), "cells unstimulated.\n"))
}
```

We have `r length(sce_unst_list)` donors for analysis.

### generate sce list with only assigned cells in clones with more than three cells

Filter donors for DE/pathway analysis using all (stimulated and unstimulated cells).

* Keep clones with at least three assigned cells.
* Keep only cells that are confidently assigned to a (kept) clone. 
* Keep donors that have at least 15 assigned cells to kept clones

Conduct the filtering for the unstimulated cells only.

```{r}
sce_unst_list_filt <- list()
clones_keep_unst <- list()
for ( i in names(sce_unst_list) ) {
    clones_keep_unst[[i]] <- table(sce_unst_list[[i]]$assigned)[which(table(sce_unst_list[[i]]$assigned) >= 3)]
    clones_keep_unst[[i]] <- clones_keep_unst[[i]][!(names(clones_keep_unst[[i]]) %in% "unassigned")]
    cells_use_unst <- colnames(sce_unst_list[[i]])[sce_unst_list[[i]]$assigned %in% names(clones_keep_unst[[i]])]
    cat(paste("....", i, ":", length(cells_use_unst), "unstimulated cells for DE analyses.\n"))
    if (length(cells_use_unst) > 14.5 && length(clones_keep_unst[[i]]) > 1.5)
        sce_unst_list_filt[[i]] <- sce_unst_list[[i]][, cells_use_unst]
}
```


`r length(sce_unst_list_filt)` donors are retained for DE/pathway analysis using
unstimulated cells only.

### filter cells and genes


### DE analysis across donors - Unstimulated cells

```{r}
data_dir <- file.path(to_working_dir, "Data/processed/de_analysis_FTv62/permutations/")
fig_dir <- file.path(to_working_dir, "figures/clonality/de_pathways/permutations/")
```

```{r}
gene_use_DE <- (!rowData(sce_unst_list_filt[[1]])$is_feature_control) # remove ERCC
for (i in names(sce_unst_list_filt)){
  gene_use_DE <- gene_use_DE & (rowMeans(counts(sce_unst_list_filt[[i]]) > 1)) # all 0s
  # 1 count in 10% of cells
  # rowMeans of counts > 0.5
}

dge_list_unst <- list()
design_list_unst <- list()
fit_list_unst <- list()
qlf_list_unst <- list()
```

```{r}
sum(gene_use_DE)
```

Create DGEList objects, calculate normalization factors, design matrix and 
estimate 

```{r}
for(i in names(sce_unst_list_filt)) {
  #sce_list_unst[[i]] <- scran::computeSumFactors(sce_list_unst[[i]])
    cat("....calculating DE for ", i, "\n")
    dge_list_unst[[i]] <- edgeR::DGEList(round(counts(sce_unst_list_filt[[i]][gene_use_DE,])))
    dge_list_unst[[i]] <- edgeR::calcNormFactors(dge_list_unst[[i]], method = "TMM")
    sce_unst_list_filt[[i]]$cdr <- (colSums(counts(sce_unst_list_filt[[i]]) > 0) / nrow(sce_unst_list_filt[[i]]))
    sce_unst_list_filt[[i]]$plate <- as.factor(sce_unst_list_filt[[i]]$plate)
    design_list_unst[[i]] <- model.matrix(~cdr + plate + assigned, data = colData(sce_unst_list_filt[[i]]))
    dge_list_unst[[i]] <- estimateDisp(dge_list_unst[[i]], design_list_unst[[i]])
 
}
```



```{r}
dge_list_unst_perm <- list()
design_list_unst_perm <- list()
fit_list_unst_perm <- list()
qlf_list_unst_perm <- list()

for(i in names(sce_unst_list_filt)) {
    design_list_unst_perm[[i]] <- list()
    fit_list_unst_perm[[i]] <- list()
    qlf_list_unst_perm[[i]] <- list()
    for (j in seq_len(nperm)) {
        sce_unst_list_filt[[i]]$assigned_perm <- 
            sample(sce_unst_list_filt[[i]]$assigned)
        design_list_unst_perm[[i]][[j]] <- model.matrix(~cdr + plate + assigned_perm, 
            data = colData(sce_unst_list_filt[[i]]))
        if (qr(design_list_unst_perm[[i]][[j]])$rank < 
            ncol(design_list_unst_perm[[i]][[j]])) {
            cat("Design matrix not of full rank for", i, "\n")
            fit_list_unst_perm[[i]][[j]] <- NA
            qlf_list_unst_perm[[i]][[j]] <- NA
        } else if (nlevels(as.factor(sce_unst_list_filt[[i]]$assigned)) < 2) {
            cat("assigned factor has only one level for ", i, "\n")
        } else {
            fit_list_unst_perm[[i]][[j]] <- glmQLFit(dge_list_unst[[i]], 
                design_list_unst_perm[[i]][[j]])
            num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
            ncoef <- ncol(design_list_unst_perm[[i]][[j]])
            qlf_list_unst_perm[[i]][[j]] <- glmQLFTest(
                fit_list_unst_perm[[i]][[j]], 
                coef = (ncoef - num_clones + 2):ncoef)
        }
    }
}
```

Conduct QL F-test for first clone coefficient alone (typically clone2 - clone1).

```{r}
qlf_1st_coef_list_unst <- list()

for(i in names(dge_list_unst)) {
    qlf_1st_coef_list_unst[[i]] <- list()
    cat("....calculating DE for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    for (j in seq_len(nperm)) {
        if (is.na(fit_list_unst_perm[[i]][[j]])) {
            qlf_1st_coef_list_unst[[i]][[j]] <- NA
        } else {
            qlf_1st_coef_list_unst[[i]][[j]] <- glmQLFTest(fit_list_unst_perm[[i]][[j]],
                coef = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 2))
        }
    }
}
```

Conduct QL F-test for second clone coefficient alone (typically clone3 - clone1).

```{r}
qlf_2nd_coef_list_unst <- list()
for(i in names(dge_list_unst)) {
    qlf_2nd_coef_list_unst[[i]] <- list()
    cat("....calculating DE for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 2.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                qlf_2nd_coef_list_unst[[i]][[j]] <- NA
            } else {
                qlf_2nd_coef_list_unst[[i]][[j]] <- glmQLFTest(fit_list_unst_perm[[i]][[j]],
                              coef = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 3))
            }
        }
    }
}
```

```{r}
qlf_3rd_coef_list_unst <- list()
for(i in names(dge_list_unst)) {
    qlf_3rd_coef_list_unst[[i]] <- list()
    cat("....calculating DE for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 3.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                qlf_3rd_coef_list_unst[[i]][[j]] <- NA
            } else {
                qlf_3rd_coef_list_unst[[i]][[j]] <- glmQLFTest(fit_list_unst_perm[[i]][[j]],
                    coef = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 4))
            }
        }
    }
}
```


### Gene set testing

```{r}
library(limma)
library(org.Hs.eg.db)
load(file.path(to_working_dir, "Data/human_c6_v5p2.rdata"))
load(file.path(to_working_dir, "Data/human_H_v5p2.rdata"))
load(file.path(to_working_dir, "Data/human_c2_v5p2.rdata"))
```

```{r}
#For the reverse map ENSEMBL2EG:
# Convert to a list
xx <- as.list(org.Hs.egENSEMBL2EG)
```

#### Oncogenic gene sets (c6)

Testing first clone coefficient (typically clone2 - clone1).

```{r}
camera_msigdb_c6_1st_coef_list_unst <- list()
fry_msigdb_c6_1st_coef_list_unst <- list()
for(i in names(sce_unst_list_filt)) {
    camera_msigdb_c6_1st_coef_list_unst[[i]] <- list()
    fry_msigdb_c6_1st_coef_list_unst[[i]] <- list()
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 1.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                next
            }
            qlf_1st_coef_list_unst[[i]][[j]]$table$ensembl_gene_id <- 
                strsplit2(rownames(qlf_1st_coef_list_unst[[i]][[j]]$table), split = "_")[,1]
            qlf_1st_coef_list_unst[[i]][[j]]$table$hgnc_symbol <- 
                strsplit2(rownames(qlf_1st_coef_list_unst[[i]][[j]]$table), split = "_")[,2]
            qlf_1st_coef_list_unst[[i]][[j]]$table$entrezid <- NA
            for (k in seq_len(nrow(qlf_1st_coef_list_unst[[i]][[j]]$table))) {
                if (qlf_1st_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                    qlf_1st_coef_list_unst[[i]][[j]]$table$entrezid[k] <- 
                        xx[[qlf_1st_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k]]][1]
            }
            idx <- ids2indices(Hs.c6, id=qlf_1st_coef_list_unst[[i]][[j]]$table$entrezid)
            camera_msigdb_c6_1st_coef_list_unst[[i]][[j]] <- camera(
                dge_list_unst[[i]], idx, qlf_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 2))
            fry_msigdb_c6_1st_coef_list_unst[[i]][[j]] <- fry(
                dge_list_unst[[i]], idx, qlf_1st_coef_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 2))
        }
    }
}
```

Testing second clone coefficient (typically clone3 - clone1).

```{r}
camera_msigdb_c6_2nd_coef_list_unst <- list()
fry_msigdb_c6_2nd_coef_list_unst <- list()
for(i in names(sce_unst_list_filt)) {
    camera_msigdb_c6_2nd_coef_list_unst[[i]] <- list()
    fry_msigdb_c6_2nd_coef_list_unst[[i]] <- list()
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 2.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                next
            }
            qlf_2nd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id <- 
                strsplit2(rownames(qlf_2nd_coef_list_unst[[i]][[j]]$table), split = "_")[,1]
            qlf_2nd_coef_list_unst[[i]][[j]]$table$hgnc_symbol <- 
                strsplit2(rownames(qlf_2nd_coef_list_unst[[i]][[j]]$table), split = "_")[,2]
            qlf_2nd_coef_list_unst[[i]][[j]]$table$entrezid <- NA
            for (k in seq_len(nrow(qlf_2nd_coef_list_unst[[i]][[j]]$table))) {
                if (qlf_2nd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                    qlf_2nd_coef_list_unst[[i]][[j]]$table$entrezid[k] <- 
                        xx[[qlf_2nd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k]]][1]
            }
            idx <- ids2indices(Hs.c6, id=qlf_2nd_coef_list_unst[[i]][[j]]$table$entrezid)
            camera_msigdb_c6_2nd_coef_list_unst[[i]][[j]] <- camera(
                dge_list_unst[[i]], idx, qlf_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 3))
            fry_msigdb_c6_2nd_coef_list_unst[[i]][[j]] <- fry(
                dge_list_unst[[i]], idx, qlf_2nd_coef_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 3))
        }
    }
}

```

Testing third clone coefficient (typically clone4 - clone1).

```{r}
camera_msigdb_c6_3rd_coef_list_unst <- list()
fry_msigdb_c6_3rd_coef_list_unst <- list()
for(i in names(sce_unst_list_filt)) {
    camera_msigdb_c6_3rd_coef_list_unst[[i]] <- list()
    fry_msigdb_c6_3rd_coef_list_unst[[i]] <- list()
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 3.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                next
            }
            qlf_3rd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id <- 
                strsplit2(rownames(qlf_3rd_coef_list_unst[[i]][[j]]$table), split = "_")[,1]
            qlf_3rd_coef_list_unst[[i]][[j]]$table$hgnc_symbol <- 
                strsplit2(rownames(qlf_3rd_coef_list_unst[[i]][[j]]$table), split = "_")[,2]
            qlf_3rd_coef_list_unst[[i]][[j]]$table$entrezid <- NA
            for (k in seq_len(nrow(qlf_3rd_coef_list_unst[[i]][[j]]$table))) {
                if (qlf_3rd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                    qlf_3rd_coef_list_unst[[i]][[j]]$table$entrezid[k] <- 
                        xx[[qlf_3rd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k]]][1]
            }
            idx <- ids2indices(Hs.c6, id=qlf_3rd_coef_list_unst[[i]][[j]]$table$entrezid)
            camera_msigdb_c6_3rd_coef_list_unst[[i]][[j]] <- camera(
                dge_list_unst[[i]], idx, qlf_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 4))
            fry_msigdb_c6_3rd_coef_list_unst[[i]][[j]] <- fry(
                dge_list_unst[[i]], idx, qlf_3rd_coef_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 4))
        }
    }
}

```

#### Curated gene sets (c2)

Testing first clone coefficient (typically clone2 - clone1).

```{r}
camera_msigdb_c2_1st_coef_list_unst <- list()
fry_msigdb_c2_1st_coef_list_unst <- list()
for(i in names(sce_unst_list_filt)) {
    camera_msigdb_c2_1st_coef_list_unst[[i]] <- list()
    fry_msigdb_c2_1st_coef_list_unst[[i]] <- list()
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 1.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                next
            }
            qlf_1st_coef_list_unst[[i]][[j]]$table$ensembl_gene_id <- 
                strsplit2(rownames(qlf_1st_coef_list_unst[[i]][[j]]$table), split = "_")[,1]
            qlf_1st_coef_list_unst[[i]][[j]]$table$hgnc_symbol <- 
                strsplit2(rownames(qlf_1st_coef_list_unst[[i]][[j]]$table), split = "_")[,2]
            qlf_1st_coef_list_unst[[i]][[j]]$table$entrezid <- NA
            for (k in seq_len(nrow(qlf_1st_coef_list_unst[[i]][[j]]$table))) {
                if (qlf_1st_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                    qlf_1st_coef_list_unst[[i]][[j]]$table$entrezid[k] <- 
                        xx[[qlf_1st_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k]]][1]
            }
            idx <- ids2indices(Hs.c2, id=qlf_1st_coef_list_unst[[i]][[j]]$table$entrezid)
            camera_msigdb_c2_1st_coef_list_unst[[i]][[j]] <- camera(
                dge_list_unst[[i]], idx, qlf_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 2))
            fry_msigdb_c2_1st_coef_list_unst[[i]][[j]] <- fry(
                dge_list_unst[[i]], idx, qlf_1st_coef_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 2))
        }
    }
}
```

Testing second clone coefficient (typically clone3 - clone1).

```{r}
camera_msigdb_c2_2nd_coef_list_unst <- list()
fry_msigdb_c2_2nd_coef_list_unst <- list()
for(i in names(sce_unst_list_filt)) {
    camera_msigdb_c2_2nd_coef_list_unst[[i]] <- list()
    fry_msigdb_c2_2nd_coef_list_unst[[i]] <- list()
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 2.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                next
            }
            qlf_2nd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id <- 
                strsplit2(rownames(qlf_2nd_coef_list_unst[[i]][[j]]$table), split = "_")[,1]
            qlf_2nd_coef_list_unst[[i]][[j]]$table$hgnc_symbol <- 
                strsplit2(rownames(qlf_2nd_coef_list_unst[[i]][[j]]$table), split = "_")[,2]
            qlf_2nd_coef_list_unst[[i]][[j]]$table$entrezid <- NA
            for (k in seq_len(nrow(qlf_2nd_coef_list_unst[[i]][[j]]$table))) {
                if (qlf_2nd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                    qlf_2nd_coef_list_unst[[i]][[j]]$table$entrezid[k] <- 
                        xx[[qlf_2nd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k]]][1]
            }
            idx <- ids2indices(Hs.c2, id=qlf_2nd_coef_list_unst[[i]][[j]]$table$entrezid)
            camera_msigdb_c2_2nd_coef_list_unst[[i]][[j]] <- camera(
                dge_list_unst[[i]], idx, qlf_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 3))
            fry_msigdb_c2_2nd_coef_list_unst[[i]][[j]] <- fry(
                dge_list_unst[[i]], idx, qlf_2nd_coef_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 3))
        }
    }
}

```

Testing third clone coefficient (typically clone4 - clone1).

```{r}
camera_msigdb_c2_3rd_coef_list_unst <- list()
fry_msigdb_c2_3rd_coef_list_unst <- list()
for(i in names(sce_unst_list_filt)) {
    camera_msigdb_c2_3rd_coef_list_unst[[i]] <- list()
    fry_msigdb_c2_3rd_coef_list_unst[[i]] <- list()
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 3.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                next
            }
            qlf_3rd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id <- 
                strsplit2(rownames(qlf_3rd_coef_list_unst[[i]][[j]]$table), split = "_")[,1]
            qlf_3rd_coef_list_unst[[i]][[j]]$table$hgnc_symbol <- 
                strsplit2(rownames(qlf_3rd_coef_list_unst[[i]][[j]]$table), split = "_")[,2]
            qlf_3rd_coef_list_unst[[i]][[j]]$table$entrezid <- NA
            for (k in seq_len(nrow(qlf_3rd_coef_list_unst[[i]][[j]]$table))) {
                if (qlf_3rd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                    qlf_3rd_coef_list_unst[[i]][[j]]$table$entrezid[k] <- 
                        xx[[qlf_3rd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k]]][1]
            }
            idx <- ids2indices(Hs.c2, id=qlf_3rd_coef_list_unst[[i]][[j]]$table$entrezid)
            camera_msigdb_c2_3rd_coef_list_unst[[i]][[j]] <- camera(
                dge_list_unst[[i]], idx, qlf_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 4))
            fry_msigdb_c2_3rd_coef_list_unst[[i]][[j]] <- fry(
                dge_list_unst[[i]], idx, qlf_3rd_coef_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 4))
        }
    }
}
```


#### Hallmark gene sets (H)

Testing first clone coefficient (typically clone2 - clone1).

```{r}
camera_msigdb_H_1st_coef_list_unst <- list()
fry_msigdb_H_1st_coef_list_unst <- list()
for(i in names(sce_unst_list_filt)) {
    camera_msigdb_H_1st_coef_list_unst[[i]] <- list()
    fry_msigdb_H_1st_coef_list_unst[[i]] <- list()
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 1.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                next
            }
            qlf_1st_coef_list_unst[[i]][[j]]$table$ensembl_gene_id <- 
                strsplit2(rownames(qlf_1st_coef_list_unst[[i]][[j]]$table), split = "_")[,1]
            qlf_1st_coef_list_unst[[i]][[j]]$table$hgnc_symbol <- 
                strsplit2(rownames(qlf_1st_coef_list_unst[[i]][[j]]$table), split = "_")[,2]
            qlf_1st_coef_list_unst[[i]][[j]]$table$entrezid <- NA
            for (k in seq_len(nrow(qlf_1st_coef_list_unst[[i]][[j]]$table))) {
                if (qlf_1st_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                    qlf_1st_coef_list_unst[[i]][[j]]$table$entrezid[k] <- 
                        xx[[qlf_1st_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k]]][1]
            }
            idx <- ids2indices(Hs.H, id=qlf_1st_coef_list_unst[[i]][[j]]$table$entrezid)
            camera_msigdb_H_1st_coef_list_unst[[i]][[j]] <- camera(
                dge_list_unst[[i]], idx, qlf_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 2))
            fry_msigdb_H_1st_coef_list_unst[[i]][[j]] <- fry(
                dge_list_unst[[i]], idx, qlf_1st_coef_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 2))
        }
    }
}
```

Testing second clone coefficient (typically clone3 - clone1).

```{r}
camera_msigdb_H_2nd_coef_list_unst <- list()
fry_msigdb_H_2nd_coef_list_unst <- list()
for(i in names(sce_unst_list_filt)) {
    camera_msigdb_H_2nd_coef_list_unst[[i]] <- list()
    fry_msigdb_H_2nd_coef_list_unst[[i]] <- list()
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 2.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                next
            }
            qlf_2nd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id <- 
                strsplit2(rownames(qlf_2nd_coef_list_unst[[i]][[j]]$table), split = "_")[,1]
            qlf_2nd_coef_list_unst[[i]][[j]]$table$hgnc_symbol <- 
                strsplit2(rownames(qlf_2nd_coef_list_unst[[i]][[j]]$table), split = "_")[,2]
            qlf_2nd_coef_list_unst[[i]][[j]]$table$entrezid <- NA
            for (k in seq_len(nrow(qlf_2nd_coef_list_unst[[i]][[j]]$table))) {
                if (qlf_2nd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                    qlf_2nd_coef_list_unst[[i]][[j]]$table$entrezid[k] <- 
                        xx[[qlf_2nd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k]]][1]
            }
            idx <- ids2indices(Hs.H, id=qlf_2nd_coef_list_unst[[i]][[j]]$table$entrezid)
            camera_msigdb_H_2nd_coef_list_unst[[i]][[j]] <- camera(
                dge_list_unst[[i]], idx, qlf_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 3))
            fry_msigdb_H_2nd_coef_list_unst[[i]][[j]] <- fry(
                dge_list_unst[[i]], idx, qlf_2nd_coef_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 3))
        }
    }
}

```

Testing third clone coefficient (typically clone4 - clone1).

```{r}
camera_msigdb_H_3rd_coef_list_unst <- list()
fry_msigdb_H_3rd_coef_list_unst <- list()
for(i in names(sce_unst_list_filt)) {
    camera_msigdb_H_3rd_coef_list_unst[[i]] <- list()
    fry_msigdb_H_3rd_coef_list_unst[[i]] <- list()
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_unst_list_filt[[i]]$assigned))
    if (num_clones > 3.5) {
        for (j in seq_len(nperm)) {
            if (is.na(fit_list_unst_perm[[i]][[j]])) {
                next
            }
            qlf_3rd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id <- 
                strsplit2(rownames(qlf_3rd_coef_list_unst[[i]][[j]]$table), split = "_")[,1]
            qlf_3rd_coef_list_unst[[i]][[j]]$table$hgnc_symbol <- 
                strsplit2(rownames(qlf_3rd_coef_list_unst[[i]][[j]]$table), split = "_")[,2]
            qlf_3rd_coef_list_unst[[i]][[j]]$table$entrezid <- NA
            for (k in seq_len(nrow(qlf_3rd_coef_list_unst[[i]][[j]]$table))) {
                if (qlf_3rd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k] %in% names(xx))
                    qlf_3rd_coef_list_unst[[i]][[j]]$table$entrezid[k] <- 
                        xx[[qlf_3rd_coef_list_unst[[i]][[j]]$table$ensembl_gene_id[k]]][1]
            }
            idx <- ids2indices(Hs.H, id=qlf_3rd_coef_list_unst[[i]][[j]]$table$entrezid)
            camera_msigdb_H_3rd_coef_list_unst[[i]][[j]] <- camera(
                dge_list_unst[[i]], idx, qlf_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 4))
            fry_msigdb_H_3rd_coef_list_unst[[i]][[j]] <- fry(
                dge_list_unst[[i]], idx, qlf_3rd_coef_list_unst[[i]][[j]]$design, 
                contrast = (ncol(design_list_unst_perm[[i]][[j]]) - num_clones + 4))
        }
    }
}
```



# Save results to file

```{r}
de_results_list_unst <- list()
```

```{r}
de_results_list_unst[["camera"]] <- list()
de_results_list_unst[["camera"]][["c2"]] <- list()
de_results_list_unst[["camera"]][["c6"]] <- list()
de_results_list_unst[["camera"]][["H"]] <- list()
```

```{r}
de_results_list_unst[["camera"]][["c2"]][["1st_coef"]] <- camera_msigdb_c2_1st_coef_list_unst
de_results_list_unst[["camera"]][["c2"]][["2nd_coef"]] <- camera_msigdb_c2_2nd_coef_list_unst
de_results_list_unst[["camera"]][["c2"]][["3rd_coef"]] <- camera_msigdb_c2_3rd_coef_list_unst
de_results_list_unst[["camera"]][["c6"]][["1st_coef"]] <- camera_msigdb_c6_1st_coef_list_unst
de_results_list_unst[["camera"]][["c6"]][["2nd_coef"]] <- camera_msigdb_c6_2nd_coef_list_unst
de_results_list_unst[["camera"]][["c6"]][["3rd_coef"]] <- camera_msigdb_c6_3rd_coef_list_unst
de_results_list_unst[["camera"]][["H"]][["1st_coef"]] <- camera_msigdb_H_1st_coef_list_unst
de_results_list_unst[["camera"]][["H"]][["2nd_coef"]] <- camera_msigdb_H_2nd_coef_list_unst
de_results_list_unst[["camera"]][["H"]][["3rd_coef"]] <- camera_msigdb_H_3rd_coef_list_unst
```

```{r}
de_results_list_unst[["fry"]] <- list()
de_results_list_unst[["fry"]][["c2"]] <- list()
de_results_list_unst[["fry"]][["c6"]] <- list()
de_results_list_unst[["fry"]][["H"]] <- list()
de_results_list_unst[["fry"]][["c2"]][["1st_coef"]] <- fry_msigdb_c2_1st_coef_list_unst
de_results_list_unst[["fry"]][["c2"]][["2nd_coef"]] <- fry_msigdb_c2_2nd_coef_list_unst
de_results_list_unst[["fry"]][["c2"]][["3rd_coef"]] <- fry_msigdb_c2_3rd_coef_list_unst
de_results_list_unst[["fry"]][["c6"]][["1st_coef"]] <- fry_msigdb_c6_1st_coef_list_unst
de_results_list_unst[["fry"]][["c6"]][["2nd_coef"]] <- fry_msigdb_c6_2nd_coef_list_unst
de_results_list_unst[["fry"]][["c6"]][["3rd_coef"]] <- fry_msigdb_c6_3rd_coef_list_unst
de_results_list_unst[["fry"]][["H"]][["1st_coef"]] <- fry_msigdb_H_1st_coef_list_unst
de_results_list_unst[["fry"]][["H"]][["2nd_coef"]] <- fry_msigdb_H_2nd_coef_list_unst
de_results_list_unst[["fry"]][["H"]][["3rd_coef"]] <- fry_msigdb_H_3rd_coef_list_unst
```

```{r}
de_results_list_unst[["design_list"]] <- design_list_unst_perm
de_results_list_unst[["dge_list"]] <- dge_list_unst_perm
de_results_list_unst[["fit_list"]] <- fit_list_unst_perm
de_results_list_unst[["qlf_list"]] <- qlf_list_unst_perm
de_results_list_unst[["qlf_1st_coef_list"]] <- qlf_1st_coef_list_unst
de_results_list_unst[["qlf_2nd_coef_list"]] <- qlf_2nd_coef_list_unst
de_results_list_unst[["qlf_3rd_coef_list"]] <- qlf_3rd_coef_list_unst
de_results_list_unst[["sce_list_unst"]] <- sce_unst_list_filt
```


```{r}
saveRDS(object = de_results_list_unst, 
    file = file.path(data_dir, 
    paste0(params$callset, ".de_results_unstimulated_cells.rds")))
```

```{r}
pryr::object_size(de_results_list_unst)
```

# Session Info

```{r}
sessionInfo()
```

