---
title: "`r params$title`"
author: "`r params$author`"
date: "`r Sys.Date()`"
params:
    callset: "MISSING"
    author: "Davis McCarthy"
    title: "Differential gene and pathway analysis between clones"
    to_working_dir: "./"

output: 
    html_document:
        toc: true
        toc_float: true
        theme: journal
        highlight: pygments
        number_sections: true
        code_folding: hide

---

# Differential gene and pathway expression analysis between clones: all cells

**Cardelino cell assignments using Canopy trees**. No merging of clones.

Using the latest callset from Petr Danecek (April 2018): fisher exact test and 
other filters. This analysis uses the callset: `r params$callset`.

```{r global_options, include=FALSE}
library(vcfR)
library(readr)
library(dplyr)
library(scran)
library(scater)
library(viridis)
library(ggplot2)
library(ggridges)
library(SingleCellExperiment)
library(edgeR)
library(limma)
library(org.Hs.eg.db)
library(cowplot)
library(gplots)
library(goseq)
```

## Load data

Define the donors for which we have cell assignment results to study.

```{r}
cat(getwd())
fls <- list.files(file.path(to_working_dir, "Data/processed/sces"))
fls <- fls[grepl(params$callset, fls)]
donors <- gsub(".*ce_([a-z]+)_.*", "\\1", fls)
#donors <- c('bubh', 'ciwj', 'deyz', 'diku', 'eipl', 'eofe', 'euts', 'fawm', 
#            'feec', 'fiaj', 'fikt', 'garx', 'gesg', 'gifk', 'hehd', 'heja', 
#            'hipn', 'ieki', 'jogf', 'joxm', 'laey', 'lexy', 'melw', 'miaj', 
#            'naju', 'nusw', 'oaaz', 'oaqd', 'oicx', 'oilg', 'pamv', 'pelm', 
#            'pipw', 'puie', 'qayj', 'qolg', 'qonc', 'rozh', 'sebz', 'sehl', 
#            'sohd', 'tixi', 'ualf', 'vabj', 'vass', 'vils', 'vuna', 'wahn', 
#            'wetu', 'wuye', 'xugn', 'zihe', 'zoxy')
length(donors)
```

Read in annotated SCE objects and create a list of SCE objects containing all 
cells (stimulated and unstimulated) and just unstimulated cells.

```{r}
sce_list <- list()
for (don in donors) {
    sce_list[[don]] <- readRDS(file.path(to_working_dir, "Data/processed/sces", 
        paste0("sce_", don, "_with_clone_assignments.", params$callset, ".rds")))
    cat(paste("reading", don, ":   ", ncol(sce_list[[don]]), "cells total.\n"))
}
```

We have `r length(sce_list)` donors for analysis.


## Plot overviews of assignments

Define data frames with numbers of cells assigned for different clones by donor.


```{r}
cells_per_clone_all <- list()
for (i in names(sce_list)) {
    if (ncol(sce_list[[i]]) == 0)
        cat(paste0("No cells for ", i, "\n"))
    else {
        tmp <- as_data_frame(table(sce_list[[i]]$assigned))
        colnames(tmp) <- c("clone", "n_cells")
        tmp[["donor"]] <- i
        tmp2 <- data.frame(clone = "total assigned", 
                           n_cells = sum(tmp[["n_cells"]][tmp[["clone"]] != "unassigned"]), donor = i)
        tmp <- bind_rows(tmp, tmp2)
        tmp[["assigned"]] <- tmp[["clone"]] != "unassigned"
        cells_per_clone_all[[i]] <- tmp
    }
}
df_cells_per_clone_all <- do.call("bind_rows", cells_per_clone_all)
```


```{r}
options(repr.plot.width = 9, repr.plot.height = 9)
df_cells_per_clone_all %>% dplyr::filter(clone != "unassigned") %>% 
ggplot(aes(x = n_cells, y = reorder(donor, n_cells, max), colour = clone)) +
geom_point(size = 4, alpha = 0.7) +
scale_shape_manual(values = c(4, 19)) +
scale_color_manual(values = c(magma(8)[-c(1,8)], "gray30", "black")) +
#ggthemes::scale_colour_tableau() +
theme_bw(16) + xlab("Number of cells") + ylab("Donor") +
ggtitle("All cells per clone, by donor")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.cells_per_clone.v1.png")),
        width = 7, height = 9)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.cells_per_clone.v1.pdf")),
        width = 7, height = 9)
```

```{r}
options(repr.plot.width = 9, repr.plot.height = 9)
ggplot(df_cells_per_clone_all, aes(x = n_cells, y = reorder(donor, n_cells, max), colour = clone, shape = assigned)) +
geom_point(size = 4, alpha = 0.7) +
scale_shape_manual(values = c(4, 19)) +
scale_color_manual(values = c(magma(8)[-c(1,8)], "gray30", "black")) +
#ggthemes::scale_colour_tableau() +
theme_bw(16) + xlab("Number of cells") + ylab("Donor") +
ggtitle("All cells per clone, by donor")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.cells_per_clone.v2.png")),
        width = 7, height = 9)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.cells_per_clone.v2.pdf")),
        width = 7, height = 9)
```


```{r}
df_cells_per_clone_all %>% dplyr::select(-assigned) %>% tidyr::spread(key = "clone", value = "n_cells")
```


### generate sce list with only assigned cells in clones with more than three cells

Filter donors for DE/pathway analysis using all (stimulated and unstimulated cells).

* Keep clones with at least three assigned cells.
* Keep only cells that are confidently assigned to a (kept) clone. 
* Keep donors that have at least 15 assigned cells to kept clones

```{r}
sce_list_filt <- list()
clones_keep <- list()
for ( i in names(sce_list) ) {
    clones_keep[[i]] <- table(sce_list[[i]]$assigned)[which(table(sce_list[[i]]$assigned) >= 3)]
    clones_keep[[i]] <- clones_keep[[i]][!(names(clones_keep[[i]]) %in% "unassigned")]
    cells_use <- colnames(sce_list[[i]])[sce_list[[i]]$assigned %in% names(clones_keep[[i]])]
    cat(paste("....", i, ":", length(cells_use), "cells for DE analyses.\n"))
    if (length(cells_use) > 14.5)
        sce_list_filt[[i]] <- sce_list[[i]][, cells_use]
}
```


```{r}
length(sce_list_filt)
```

`r length(sce_list_filt)` donors are retained for DE/pathway analyses from all cells.



## DE analysis across donors - all cells (stimulated and unstimulated)


```{r}
gene_use_DE <- (!rowData(sce_list_filt[[1]])$is_feature_control) # remove ERCC
for (i in names(sce_list_filt)){
  gene_use_DE <- gene_use_DE & (rowMeans(counts(sce_list_filt[[i]]) > 1)) # all 0s
  # 1 count in 10% of cells
  # rowMeans of counts > 0.5
}

dge_list_all <- list()
design_list_all <- list()
fit_list_all <- list()
qlf_list_all <- list()
```

```{r}
sum(gene_use_DE)
```

```{r}
length(sce_list_filt)
```

```{r}
table(sce_list_filt[["diku"]]$plate, sce_list_filt[["diku"]]$well_condition)
```

```{r}
table(sce_list_filt[["vass"]]$plate, sce_list_filt[["vass"]]$well_condition)
```

```{r}
for(i in names(sce_list_filt)) {
    cat("....calculating DE for ", i, "\n")
    dge_list_all[[i]] <- DGEList(round(counts(sce_list_filt[[i]][gene_use_DE,])))
    dge_list_all[[i]] <- edgeR::calcNormFactors(dge_list_all[[i]], method = "TMM")
    sce_list_filt[[i]]$cdr <- (colSums(counts(sce_list_filt[[i]]) > 0) / nrow(sce_list_filt[[i]]))
    sce_list_filt[[i]]$plate <- as.factor(sce_list_filt[[i]]$plate)
    sce_list_filt[[i]]$well_condition <- as.factor(sce_list_filt[[i]]$well_condition)
    design_list_all[[i]] <- model.matrix(~cdr + plate + well_condition + assigned, 
                                            data = colData(sce_list_filt[[i]]))
    if (qr(design_list_all[[i]])$rank < ncol(design_list_all[[i]])) {
        cat("Design matrix not of full rank for", i, "\n")
    } else if (nlevels(as.factor(sce_list_filt[[i]]$assigned)) < 2) {
        cat("assigned factor has only one level for ", i, "\n")
    } else {
        dge_list_all[[i]] <- estimateDisp(dge_list_all[[i]], design_list_all[[i]])
        fit_list_all[[i]] <- glmQLFit(dge_list_all[[i]], design_list_all[[i]])
        num_clones <- length(unique(sce_list_filt[[i]]$assigned))
        qlf_list_all[[i]] <- glmQLFTest(fit_list_all[[i]],
                              coef = (ncol(design_list_all[[i]]) - num_clones + 2):ncol(design_list_all[[i]]))
        sum(p.adjust(qlf_list_all[[i]]$table$PValue, method = "BH") <= 0.05, na.rm = TRUE)
    }
}
```

```{r}
sce_list_filt[[i]]
```

Conduct QL F-test for first clone coefficient alone (typically clone2 - clone1).

```{r}
qlf_1st_coef_list_all <- list()
```

```{r}
pryr::mem_used()
```

```{r}
for(i in names(fit_list_all)) {
  cat("....calculating DE for ", i, "\n")
  num_clones <- length(unique(sce_list_filt[[i]]$assigned))
  qlf_1st_coef_list_all[[i]] <- glmQLFTest(fit_list_all[[i]],
                              coef = (ncol(design_list_all[[i]]) - num_clones + 2))
  print(summary(decideTestsDGE(qlf_1st_coef_list_all[[i]])))
}
```

Conduct QL F-test for second clone coefficient alone (typically clone3 - clone1).

```{r}
qlf_2nd_coef_list_all <- list()
for(i in names(fit_list_all)) {
    cat("....calculating DE for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 2.5) {
        qlf_2nd_coef_list_all[[i]] <- glmQLFTest(fit_list_all[[i]],
                              coef = (ncol(design_list_all[[i]]) - num_clones + 3))
        print(summary(decideTestsDGE(qlf_2nd_coef_list_all[[i]])))
    }
}
```

```{r}
qlf_3rd_coef_list_all <- list()
for(i in names(fit_list_all)) {
    cat("....calculating DE for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 3.5) {
        qlf_3rd_coef_list_all[[i]] <- glmQLFTest(fit_list_all[[i]],
                              coef = (ncol(design_list_all[[i]]) - num_clones + 4))
        print(summary(decideTestsDGE(qlf_3rd_coef_list_all[[i]])))
    }
}
```

```{r}
pryr::mem_used()
```

### Gene set testing

```{r}

load(file.path(to_working_dir, "Data/human_c6_v5p2.rdata"))
load(file.path(to_working_dir, "Data/human_H_v5p2.rdata"))
load(file.path(to_working_dir, "Data/human_c2_v5p2.rdata"))
#For the reverse map ENSEMBL2EG:
# Convert to a list
xx <- as.list(org.Hs.egENSEMBL2EG)
```

#### Oncogenic gene sets (c6)

Testing first clone coefficient (typically clone2 - clone1).

```{r}
camera_msigdb_c6_1st_coef_list_all <- list()
fry_msigdb_c6_1st_coef_list_all <- list()
for(i in names(fit_list_all)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 1.5) {
        qlf_1st_coef_list_all[[i]]$table$ensembl_gene_id <- strsplit2(rownames(qlf_1st_coef_list_all[[i]]$table), split = "_")[,1]
        qlf_1st_coef_list_all[[i]]$table$hgnc_symbol <- strsplit2(rownames(qlf_1st_coef_list_all[[i]]$table), split = "_")[,2]
        qlf_1st_coef_list_all[[i]]$table$entrezid <- NA
        for (j in seq_len(nrow(qlf_1st_coef_list_all[[i]]$table))) {
            if (qlf_1st_coef_list_all[[i]]$table$ensembl_gene_id[j] %in% names(xx))
                qlf_1st_coef_list_all[[i]]$table$entrezid[j] <- xx[[qlf_1st_coef_list_all[[i]]$table$ensembl_gene_id[j]]][1]
        }
        idx <- ids2indices(Hs.c6, id=qlf_1st_coef_list_all[[i]]$table$entrezid)
        camera_msigdb_c6_1st_coef_list_all[[i]] <- camera(dge_list_all[[i]], idx, qlf_list_all[[i]]$design, 
                                                      contrast = (ncol(design_list_all[[i]]) - num_clones + 2))
        cat("                camera significant MSigDB oncogenic genesets (FDR < 5%): ", 
           sum(camera_msigdb_c6_1st_coef_list_all[[i]]$FDR < 0.05), "\n")
        fry_msigdb_c6_1st_coef_list_all[[i]] <- fry(dge_list_all[[i]], idx, qlf_1st_coef_list_all[[i]]$design, 
                                                contrast = (ncol(design_list_all[[i]]) - num_clones + 2))    
        cat("                fry significant MSigDB oncogenic genesets (FDR < 5%): ", 
           sum(fry_msigdb_c6_1st_coef_list_all[[i]]$FDR < 0.05), "\n")
    }
}
```

Testing second clone coefficient (typically clone3 - clone1).

```{r}
camera_msigdb_c6_2nd_coef_list_all <- list()
fry_msigdb_c6_2nd_coef_list_all <- list()
for(i in names(fit_list_all)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 2.5) {
        qlf_2nd_coef_list_all[[i]]$table$ensembl_gene_id <- strsplit2(rownames(qlf_2nd_coef_list_all[[i]]$table), split = "_")[,1]
        qlf_2nd_coef_list_all[[i]]$table$hgnc_symbol <- strsplit2(rownames(qlf_2nd_coef_list_all[[i]]$table), split = "_")[,2]
        qlf_2nd_coef_list_all[[i]]$table$entrezid <- NA
        for (j in seq_len(nrow(qlf_2nd_coef_list_all[[i]]$table))) {
            if (qlf_2nd_coef_list_all[[i]]$table$ensembl_gene_id[j] %in% names(xx))
                qlf_2nd_coef_list_all[[i]]$table$entrezid[j] <- xx[[qlf_2nd_coef_list_all[[i]]$table$ensembl_gene_id[j]]][1]
        }
        idx <- ids2indices(Hs.c6, id=qlf_2nd_coef_list_all[[i]]$table$entrezid)
        camera_msigdb_c6_2nd_coef_list_all[[i]] <- camera(dge_list_all[[i]], idx, qlf_list_all[[i]]$design, 
                                                      contrast = (ncol(design_list_all[[i]]) - num_clones + 3))
        cat("                camera significant MSigDB oncogenic genesets (FDR < 5%): ", 
           sum(camera_msigdb_c6_2nd_coef_list_all[[i]]$FDR < 0.05), "\n")
        fry_msigdb_c6_2nd_coef_list_all[[i]] <- fry(dge_list_all[[i]], idx, qlf_2nd_coef_list_all[[i]]$design, 
                                                contrast = (ncol(design_list_all[[i]]) - num_clones + 3))    
        cat("                fry significant MSigDB oncogenic genesets (FDR < 5%): ", 
           sum(fry_msigdb_c6_2nd_coef_list_all[[i]]$FDR < 0.05), "\n")
    }
}
```

Testing third clone coefficient (typically clone4 - clone1).

```{r}
camera_msigdb_c6_3rd_coef_list_all <- list()
fry_msigdb_c6_3rd_coef_list_all <- list()
for(i in names(fit_list_all)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 3.5) {
        qlf_3rd_coef_list_all[[i]]$table$ensembl_gene_id <- strsplit2(rownames(qlf_3rd_coef_list_all[[i]]$table), split = "_")[,1]
        qlf_3rd_coef_list_all[[i]]$table$hgnc_symbol <- strsplit2(rownames(qlf_3rd_coef_list_all[[i]]$table), split = "_")[,2]
        qlf_3rd_coef_list_all[[i]]$table$entrezid <- NA
        for (j in seq_len(nrow(qlf_3rd_coef_list_all[[i]]$table))) {
            if (qlf_3rd_coef_list_all[[i]]$table$ensembl_gene_id[j] %in% names(xx))
                qlf_3rd_coef_list_all[[i]]$table$entrezid[j] <- xx[[qlf_3rd_coef_list_all[[i]]$table$ensembl_gene_id[j]]][1]
        }
        idx <- ids2indices(Hs.c6, id=qlf_3rd_coef_list_all[[i]]$table$entrezid)
        camera_msigdb_c6_3rd_coef_list_all[[i]] <- camera(dge_list_all[[i]], idx, qlf_list_all[[i]]$design, 
                                                      contrast = (ncol(design_list_all[[i]]) - num_clones + 4))
        cat("                camera significant MSigDB hallmark genesets (FDR < 5%): ", 
           sum(camera_msigdb_c6_3rd_coef_list_all[[i]]$FDR < 0.05), "\n")
        fry_msigdb_c6_3rd_coef_list_all[[i]] <- fry(dge_list_all[[i]], idx, qlf_3rd_coef_list_all[[i]]$design, 
                                                contrast = (ncol(design_list_all[[i]]) - num_clones + 4))    
        cat("                fry significant MSigDB hallmark genesets (FDR < 5%): ", 
           sum(fry_msigdb_c6_3rd_coef_list_all[[i]]$FDR < 0.05), "\n")
    }
}
```

#### Curated gene sets (c2)

Testing first clone coefficient (typically clone2 - clone1).

```{r}
camera_msigdb_c2_1st_coef_list_all <- list()
fry_msigdb_c2_1st_coef_list_all <- list()
for (i in names(fit_list_all)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 1.5) {
        idx <- ids2indices(Hs.c2, id=qlf_1st_coef_list_all[[i]]$table$entrezid)
        camera_msigdb_c2_1st_coef_list_all[[i]] <- camera(dge_list_all[[i]], idx, qlf_list_all[[i]]$design, 
                                                      contrast = (ncol(design_list_all[[i]]) - num_clones + 2))
        cat("                camera significant MSigDB c2 genesets (FDR < 5%): ", 
           sum(camera_msigdb_c2_1st_coef_list_all[[i]]$FDR < 0.05), "\n")
        fry_msigdb_c2_1st_coef_list_all[[i]] <- fry(dge_list_all[[i]], idx, qlf_1st_coef_list_all[[i]]$design, 
                                                contrast = (ncol(design_list_all[[i]]) - num_clones + 2))    
        cat("                fry significant MSigDB c2 genesets (FDR < 5%): ", 
           sum(fry_msigdb_c2_1st_coef_list_all[[i]]$FDR < 0.05), "\n")
    }
}
```

Testing second clone coefficient (typically clone3 - clone1).

```{r}
camera_msigdb_c2_2nd_coef_list_all <- list()
fry_msigdb_c2_2nd_coef_list_all <- list()
for (i in names(fit_list_all)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 2.5) {
        idx <- ids2indices(Hs.c2, id=qlf_2nd_coef_list_all[[i]]$table$entrezid)
        camera_msigdb_c2_2nd_coef_list_all[[i]] <- camera(dge_list_all[[i]], idx, qlf_list_all[[i]]$design, 
                                                      contrast = (ncol(design_list_all[[i]]) - num_clones + 3))
        cat("                camera significant MSigDB c2 genesets (FDR < 5%): ", 
           sum(camera_msigdb_c2_2nd_coef_list_all[[i]]$FDR < 0.05), "\n")
        fry_msigdb_c2_2nd_coef_list_all[[i]] <- fry(dge_list_all[[i]], idx, qlf_2nd_coef_list_all[[i]]$design, 
                                                contrast = (ncol(design_list_all[[i]]) - num_clones + 3))    
        cat("                fry significant MSigDB c2 genesets (FDR < 5%): ", 
           sum(fry_msigdb_c2_2nd_coef_list_all[[i]]$FDR < 0.05), "\n")
    }
}
```

Testing third clone coefficient (typically clone4 - clone1).

```{r}
camera_msigdb_c2_3rd_coef_list_all <- list()
fry_msigdb_c2_3rd_coef_list_all <- list()
for (i in names(fit_list_all)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 3.5) {
        idx <- ids2indices(Hs.c2, id=qlf_3rd_coef_list_all[[i]]$table$entrezid)
        camera_msigdb_c2_3rd_coef_list_all[[i]] <- camera(dge_list_all[[i]], idx, qlf_list_all[[i]]$design, 
                                                      contrast = (ncol(design_list_all[[i]]) - num_clones + 4))
        cat("                camera significant MSigDB c2 genesets (FDR < 5%): ", 
           sum(camera_msigdb_c2_3rd_coef_list_all[[i]]$FDR < 0.05), "\n")
        fry_msigdb_c2_3rd_coef_list_all[[i]] <- fry(dge_list_all[[i]], idx, qlf_3rd_coef_list_all[[i]]$design, 
                                                contrast = (ncol(design_list_all[[i]]) - num_clones + 4))    
        cat("                fry significant MSigDB c2 genesets (FDR < 5%): ", 
           sum(fry_msigdb_c2_3rd_coef_list_all[[i]]$FDR < 0.05), "\n")
    }
}
```

#### Hallmark gene sets (H)

Testing first clone coefficient (typically clone2 - clone1).

```{r}
camera_msigdb_H_1st_coef_list_all <- list()
fry_msigdb_H_1st_coef_list_all <- list()
for (i in names(fit_list_all)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 1.5) {
        idx <- ids2indices(Hs.H, id=qlf_1st_coef_list_all[[i]]$table$entrezid)
        camera_msigdb_H_1st_coef_list_all[[i]] <- camera(dge_list_all[[i]], idx, qlf_list_all[[i]]$design, 
                                                      contrast = (ncol(design_list_all[[i]]) - num_clones + 2))
        cat("                camera significant MSigDB hallmark genesets (FDR < 5%): ", 
           sum(camera_msigdb_H_1st_coef_list_all[[i]]$FDR < 0.05), "\n")
        fry_msigdb_H_1st_coef_list_all[[i]] <- fry(dge_list_all[[i]], idx, qlf_1st_coef_list_all[[i]]$design, 
                                                contrast = (ncol(design_list_all[[i]]) - num_clones + 2))    
        cat("                fry significant MSigDB hallmark genesets (FDR < 5%): ", 
           sum(fry_msigdb_H_1st_coef_list_all[[i]]$FDR < 0.05), "\n")
    }
}
```

Testing second clone coefficient (typically clone3 - clone1).

```{r}
camera_msigdb_H_2nd_coef_list_all <- list()
fry_msigdb_H_2nd_coef_list_all <- list()
for(i in names(fit_list_all)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 2.5) {
        idx <- ids2indices(Hs.H, id=qlf_2nd_coef_list_all[[i]]$table$entrezid)
        camera_msigdb_H_2nd_coef_list_all[[i]] <- camera(dge_list_all[[i]], idx, qlf_list_all[[i]]$design, 
                                                      contrast = (ncol(design_list_all[[i]]) - num_clones + 3))
        cat("                camera significant MSigDB hallmark genesets (FDR < 5%): ", 
           sum(camera_msigdb_H_2nd_coef_list_all[[i]]$FDR < 0.05), "\n")
        fry_msigdb_H_2nd_coef_list_all[[i]] <- fry(dge_list_all[[i]], idx, qlf_2nd_coef_list_all[[i]]$design, 
                                                contrast = (ncol(design_list_all[[i]]) - num_clones + 3))    
        cat("                fry significant MSigDB hallmark genesets (FDR < 5%): ", 
           sum(fry_msigdb_H_2nd_coef_list_all[[i]]$FDR < 0.05), "\n")
    }
}
```

Testing third clone coefficient (typically clone4 - clone1).

```{r}
camera_msigdb_H_3rd_coef_list_all <- list()
fry_msigdb_H_3rd_coef_list_all <- list()
for(i in names(fit_list_all)) {
    cat("....calculating gene set enrichment for ", i, "\n")
    num_clones <- length(unique(sce_list_filt[[i]]$assigned))
    if (num_clones > 3.5) {
        idx <- ids2indices(Hs.H, id=qlf_3rd_coef_list_all[[i]]$table$entrezid)
        camera_msigdb_H_3rd_coef_list_all[[i]] <- camera(dge_list_all[[i]], idx, qlf_list_all[[i]]$design, 
                                                      contrast = (ncol(design_list_all[[i]]) - num_clones + 4))
        cat("                camera significant MSigDB hallmark genesets (FDR < 5%): ", 
           sum(camera_msigdb_H_3rd_coef_list_all[[i]]$FDR < 0.05), "\n")
        fry_msigdb_H_3rd_coef_list_all[[i]] <- fry(dge_list_all[[i]], idx, qlf_3rd_coef_list_all[[i]]$design, 
                                                contrast = (ncol(design_list_all[[i]]) - num_clones + 4))    
        cat("                fry significant MSigDB hallmark genesets (FDR < 5%): ", 
           sum(fry_msigdb_H_3rd_coef_list_all[[i]]$FDR < 0.05), "\n")
    }
}
```

```{r}
de_results_list_all <- list()
```

```{r}
de_results_list_all[["camera"]] <- list()
de_results_list_all[["camera"]][["c2"]] <- list()
de_results_list_all[["camera"]][["c6"]] <- list()
de_results_list_all[["camera"]][["H"]] <- list()
```

```{r}
de_results_list_all[["camera"]][["c2"]][["1st_coef"]] <- camera_msigdb_c2_1st_coef_list_all
de_results_list_all[["camera"]][["c2"]][["2nd_coef"]] <- camera_msigdb_c2_2nd_coef_list_all
de_results_list_all[["camera"]][["c2"]][["3rd_coef"]] <- camera_msigdb_c2_3rd_coef_list_all
de_results_list_all[["camera"]][["c6"]][["1st_coef"]] <- camera_msigdb_c6_1st_coef_list_all
de_results_list_all[["camera"]][["c6"]][["2nd_coef"]] <- camera_msigdb_c6_2nd_coef_list_all
de_results_list_all[["camera"]][["c6"]][["3rd_coef"]] <- camera_msigdb_c6_3rd_coef_list_all
de_results_list_all[["camera"]][["H"]][["1st_coef"]] <- camera_msigdb_H_1st_coef_list_all
de_results_list_all[["camera"]][["H"]][["2nd_coef"]] <- camera_msigdb_H_2nd_coef_list_all
de_results_list_all[["camera"]][["H"]][["3rd_coef"]] <- camera_msigdb_H_3rd_coef_list_all
```

```{r}
de_results_list_all[["fry"]] <- list()
de_results_list_all[["fry"]][["c2"]] <- list()
de_results_list_all[["fry"]][["c6"]] <- list()
de_results_list_all[["fry"]][["H"]] <- list()
de_results_list_all[["fry"]][["c2"]][["1st_coef"]] <- fry_msigdb_c2_1st_coef_list_all
de_results_list_all[["fry"]][["c2"]][["2nd_coef"]] <- fry_msigdb_c2_2nd_coef_list_all
de_results_list_all[["fry"]][["c2"]][["3rd_coef"]] <- fry_msigdb_c2_3rd_coef_list_all
de_results_list_all[["fry"]][["c6"]][["1st_coef"]] <- fry_msigdb_c6_1st_coef_list_all
de_results_list_all[["fry"]][["c6"]][["2nd_coef"]] <- fry_msigdb_c6_2nd_coef_list_all
de_results_list_all[["fry"]][["c6"]][["3rd_coef"]] <- fry_msigdb_c6_3rd_coef_list_all
de_results_list_all[["fry"]][["H"]][["1st_coef"]] <- fry_msigdb_H_1st_coef_list_all
de_results_list_all[["fry"]][["H"]][["2nd_coef"]] <- fry_msigdb_H_2nd_coef_list_all
de_results_list_all[["fry"]][["H"]][["3rd_coef"]] <- fry_msigdb_H_3rd_coef_list_all
```

```{r}
de_results_list_all[["design_list"]] <- design_list_all
de_results_list_all[["dge_list"]] <- dge_list_all
de_results_list_all[["fit_list"]] <- fit_list_all
de_results_list_all[["qlf_list"]] <- qlf_list_all
de_results_list_all[["qlf_1st_coef_list"]] <- qlf_1st_coef_list_all
de_results_list_all[["qlf_2nd_coef_list"]] <- qlf_2nd_coef_list_all
de_results_list_all[["qlf_3rd_coef_list"]] <- qlf_3rd_coef_list_all
de_results_list_all[["sce_list"]] <- sce_list_filt
```

```{r}
data_dir <- file.path(to_working_dir, "Data/processed/de_analysis_FTv62/")
fig_dir <- file.path(to_working_dir, "figures/clonality/de_pathways/")
data_dir
```

```{r}
saveRDS(object = de_results_list_all, 
    file = file.path(data_dir, 
    paste0(params$callset, ".de_results_all_cells.rds")))
```

```{r}
pryr::object_size(de_results_list_all)
```

### Camera results

```{r}
fdr_thresh <- 0.05
df_camera_sig_all <- data_frame()
for (geneset in names(de_results_list_all[["camera"]])) {
    for (coeff in names(de_results_list_all[["camera"]][[geneset]])) {
        for (donor in names(de_results_list_all[["camera"]][[geneset]][[coeff]])) {
            tmp <- de_results_list_all[["camera"]][[geneset]][[coeff]][[donor]]
            tmp <- tmp[tmp$FDR < 0.05,]
            if (nrow(tmp) > 0.5) {
                tmp[["collection"]] <- geneset
                tmp[["geneset"]] <- rownames(tmp)
                tmp[["coeff"]] <- coeff
                tmp[["donor"]] <- donor
                df_camera_sig_all <- bind_rows(df_camera_sig_all, tmp)
            }
        }
    }
}
```

```{r}
nrow(df_camera_sig_all)
```

```{r}
head(df_camera_sig_all)
```

```{r}
df_camera_sig_all <- dplyr::mutate(df_camera_sig_all,
                              contrast = plyr::mapvalues(coeff, from = c("1st_coef", "2nd_coef", "3rd_coef"), to = c("clone2 - clone1", "clone3 - clone1", "clone4 - clone1")),
                              msigdb_collection = plyr::mapvalues(collection, from = c("c2", "c6", "H"), to = c("MSigDB curated (c2)", "MSigDB oncogenic (c6)", "MSigDB Hallmark")))
```

```{r}
head(df_camera_sig_all)
```

```{r}
table(df_camera_sig_all[["contrast"]])
```

```{r}
options(repr.plot.height = 9, repr.plot.width = 12)
ggplot(df_camera_sig_all, aes(y = -log10(PValue), x = donor, colour = contrast)) +
    geom_sina(alpha = 0.7) +
    facet_wrap(~msigdb_collection, scales = "free_x") + 
    ggthemes::scale_colour_few() +
    coord_flip() + ggtitle("All cells")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.camera_pvals.png")),
        width = 12, height = 9)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.camera_pvals.pdf")),
        width = 12, height = 9)
```

```{r}
options(repr.plot.height = 10, repr.plot.width = 12)
group_by(df_camera_sig_all, donor, msigdb_collection, contrast) %>% 
    summarise(n_sig = n()) %>%
    ungroup() %>%
ggplot(aes(y = n_sig, x = reorder(donor, n_sig, max), colour = contrast)) +
    geom_point(alpha = 0.7, size = 2) +
    facet_wrap(~msigdb_collection, scales = "free_x") + 
    ggthemes::scale_colour_tableau() +
    coord_flip() +
    theme_bw(16) +
    xlab("Donor") + ylab("Number of significant genesets") +
    ggtitle("Camera gene set enrichment results - all cells")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_genesets.v1.png")),
        width = 12, height = 10)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_genesets.v1.pdf")),
        width = 12, height = 10)
```

```{r}
options(repr.plot.height = 10, repr.plot.width = 12)
group_by(df_camera_sig_all, donor, msigdb_collection) %>% 
    dplyr::mutate(id = paste0(donor, geneset)) %>% distinct(id, .keep_all = TRUE) %>%
    summarise(n_sig = n()) %>%
    ungroup() %>%
ggplot(aes(y = n_sig, x = reorder(donor, n_sig, max))) +
    geom_point(alpha = 0.7, size = 3) +
    facet_wrap(~msigdb_collection, scales = "free_x") + 
    ggthemes::scale_colour_tableau() +
    coord_flip() +
    theme_bw(16) +
    xlab("Donor") + ylab("Number of significant genesets") +
    ggtitle("Camera gene set enrichment results (all cells)")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_genesets.v2.png")),
        width = 12, height = 10)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_genesets.v2.pdf")),
        width = 12, height = 10)
```

```{r}
df_camera_sig_all %>% dplyr::filter(donor == "sohd", collection == "H") 
```

```{r}
df_camera_sig_all %>% dplyr::filter(donor == "laey", collection == "H") 
```

```{r}
df_camera_sig_all %>% dplyr::filter(donor == "wetu", collection == "H") 
```

```{r}
df_camera_sig_all %>% dplyr::filter(collection == "H") %>% dplyr::arrange(FDR) %>% head
```


```{r}
df_camera_sig_all %>% dplyr::filter(donor == "sehl", collection == "H")  %>% dplyr::arrange(FDR)
```


```{r}
options(repr.plot.width = 9, repr.plot.height = 5)
if (is.numeric(de_results_list_all$qlf_2nd_coef_list$sehl$table$logFC)) {
    idx <- ids2indices(Hs.H, id=de_results_list_all$qlf_2nd_coef_list$sehl$table$entrezid)
    barcodeplot(de_results_list_all$qlf_2nd_coef_list$sehl$table$logFC, index=idx$HALLMARK_E2F_TARGETS, 
            index2=idx$HALLMARK_G2M_CHECKPOINT, xlab = "logFC", main="sehl: Clone3 - Clone1\n HALLMARK_E2F_TARGETS and HALLMARK_G2M_CHECKPOINT")
}
```

```{r}
df_camera_sig_all %>% dplyr::filter(donor == "laey", collection == "H") %>% dplyr::arrange(FDR)
```


```{r}
options(repr.plot.width = 11, repr.plot.height = 7)
df_camera_sig_all %>% dplyr::filter(collection == "H") %>% group_by(geneset, contrast, Direction) %>% 
    summarise(n_donors = n()) %>% dplyr::arrange(geneset, n_donors) %>% ungroup() %>%
ggplot(aes(y = n_donors, x = reorder(geneset, n_donors, max), colour = Direction, shape = contrast)) +
    geom_point(alpha = 0.7, size = 4) +
    ggthemes::scale_colour_tableau() +
    coord_flip() +
    theme_bw(16) +
    xlab("Gene set") + ylab("Number of significant donors")  +
    ggtitle("Camera MSigDB Hallmark gene set enrichment")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_hallmark_genesets.v1.png")),
        width = 11, height = 7)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_hallmark_genesets.v1.pdf")),
        width = 11, height = 7)
```

```{r}
options(repr.plot.width = 11, repr.plot.height = 7.5)
df_camera_sig_all %>% dplyr::filter(collection == "H") %>% group_by(geneset) %>% 
    dplyr::mutate(id = paste0(donor, geneset)) %>% distinct(id, .keep_all = TRUE) %>%
    summarise(n_donors = n()) %>% dplyr::arrange(geneset, n_donors) %>% ungroup() %>%
ggplot(aes(y = n_donors, x = reorder(geneset, n_donors, max))) +
    geom_point(alpha = 0.7, size = 4) +
    ggthemes::scale_colour_tableau() +
    coord_flip() +
    theme_bw(16) +
    xlab("Gene set") + ylab("Number of significant donors")  +
    ggtitle("Camera MSigDB Hallmark gene set enrichment")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_hallmark_genesets.v2.png")),
        width = 11, height = 7.5)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_hallmark_genesets.v2.pdf")),
        width = 11, height = 7.5)
```

```{r}
options(repr.plot.width = 11, repr.plot.height = 7.5)
df_camera_sig_all %>% dplyr::filter(collection == "c6") %>% group_by(geneset, contrast, Direction) %>% 
    summarise(n_donors = n()) %>% dplyr::arrange(geneset, n_donors) %>% ungroup() %>%
ggplot(aes(y = n_donors, x = reorder(geneset, n_donors, max), colour = Direction, shape = contrast)) +
    geom_point(alpha = 0.7, size = 3) +
    ggthemes::scale_colour_tableau() +
    coord_flip() +
    theme_bw(16) +
    xlab("Gene set") + ylab("Number of significant donors")  +
    ggtitle("Camera MSigDB oncogenic gene set enrichment")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_oncogenic_genesets.png")),
        width = 11, height = 7.5)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_oncogenic_genesets.pdf")),
        width = 11, height = 7.5)
```

```{r}
options(repr.plot.width = 12, repr.plot.height = 14)
df_camera_sig_all %>% dplyr::filter(collection == "c2") %>% group_by(geneset, contrast, Direction) %>% 
    summarise(n_donors = n()) %>% dplyr::arrange(geneset, n_donors) %>% ungroup() %>%
    dplyr::filter(geneset %in% geneset[n_donors > 6.5]) %>% dplyr::mutate(geneset_short = substr(geneset, 1, 40)) %>%
ggplot(aes(y = n_donors, x = reorder(geneset_short, n_donors, max), colour = Direction, shape = contrast)) +
    geom_point(alpha = 0.7, size = 4) +
    ggthemes::scale_colour_tableau() +
    coord_flip() +
    theme_bw(14) +
    xlab("Gene set") + ylab("Number of significant donors")  +
    ggtitle("Camera MSigDB curated (c2) gene set enrichment")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_c2_genesets.v1.png")),
        width = 12, height = 14)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_c2_genesets.v1.pdf")),
        width = 12, height = 14)
```

```{r}
options(repr.plot.width = 12, repr.plot.height = 14)
df_camera_sig_all %>% dplyr::filter(collection == "c2") %>% group_by(geneset, contrast, Direction) %>% 
    summarise(n_donors = n()) %>% dplyr::arrange(geneset, n_donors) %>% ungroup() %>%
    dplyr::filter(geneset %in% geneset[n_donors > 6.5]) %>% dplyr::mutate(geneset_short = substr(geneset, 1, 40)) %>%
ggplot(aes(y = n_donors, x = reorder(geneset_short, n_donors, max), colour = contrast, shape = contrast)) +
    geom_point(alpha = 0.7, size = 4) +
    ggthemes::scale_colour_tableau() +
    coord_flip() +
    theme_bw(14) +
    xlab("Gene set") + ylab("Number of significant donors")  +
    ggtitle("Camera MSigDB curated (c2) gene set enrichment")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_c2_genesets.v2.png")),
        width = 12, height = 14)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_c2_genesets.v2.pdf")),
        width = 12, height = 14)
```

```{r, eval = FALSE, include = FALSE}
options(repr.plot.width = 12, repr.plot.height = 14)
df_camera_sig_all %>% dplyr::filter(collection == "c2") %>% group_by(geneset) %>% 
    dplyr::mutate(id = paste0(donor, geneset)) %>% distinct(id, .keep_all = TRUE) %>%
    summarise(n_donors = n()) %>% dplyr::arrange(geneset, n_donors) %>% ungroup() %>%
    dplyr::filter(geneset %in% geneset[n_donors > 19.5]) %>% dplyr::mutate(geneset_short = substr(geneset, 1, 40)) %>%
ggplot(aes(y = n_donors, x = reorder(geneset_short, n_donors, max))) +
    geom_segment(aes(x = reorder(geneset_short, n_donors, max), 
    xend = reorder(geneset_short, n_donors, max), y = 0, yend = n_donors),
                alpha = 0.5) +
    geom_point(colour = "gray40", size = 4) +
    ggthemes::scale_colour_tableau() +
    coord_flip() +
    theme_classic(14) + ylim(0, 31) +
    xlab("Gene set") + ylab("Number of significant donors")  +
    ggtitle("Camera MSigDB curated (c2) gene set enrichment")

ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_c2_genesets.v3.png")),
        width = 12, height = 14)
ggsave(file.path(to_working_dir, "figures/clonality/de_pathways/", 
        paste0(params$callset, ".all_cells.nsig_c2_genesets.v3.pdf")),
        width = 12, height = 14)
```

