---
title: "Selection Models"
author: "Daniel J. Kunz & Davis J. McCarthy"
site: workflowr::wflow_site
---

## Load libraries and data

```{r setup, include=TRUE, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
library(ggrepel)
library(neutralitytestr)
library(cowplot)
library(plyr)
dir.create("figures/selection", showWarnings = FALSE, recursive = TRUE)
```

Load the call set and extract the allele frequencies which used for the fits of the selection models.

```{r}
filteredAF = read.table("data/exome-point-mutations/high-vs-low-exomes.v62.ft.filt_lenient-alldonors.txt.gz", 
                        header = TRUE, stringsAsFactors = FALSE)

mut_list = data.frame("sampleID" = filteredAF$donor_short_id,
                      "af_fibro" = filteredAF$nALT_fibro/(filteredAF$nREF_fibro + filteredAF$nALT_fibro),
                      "af_ips" = filteredAF$nALT_ips/(filteredAF$nREF_ips + filteredAF$nALT_ips),
                      "chr" = filteredAF$chrom,
                      "pos" = filteredAF$pos,
                      "ref" = filteredAF$ref,
                      "mut" = filteredAF$alt,
                      "mutID" = paste(filteredAF$chrom, filteredAF$pos, filteredAF$ref, filteredAF$alt, sep = "_"))

mut_list = mut_list[order(mut_list$sampleID),]
write.table(mut_list, "data/selection/ips-fibro-AF.tsv", 
            row.names = FALSE, quote = FALSE, sep = "\t")

mut_list = data.frame("sampleID" = filteredAF$donor_short_id,
                      "af" = filteredAF$nALT_fibro/(filteredAF$nREF_fibro + filteredAF$nALT_fibro),
                      "chr" = filteredAF$chrom,
                      "pos" = filteredAF$pos,
                      "ref" = filteredAF$ref,
                      "mut" = filteredAF$alt)
mut_list = mut_list[order(mut_list$sampleID),]

write.table(mut_list, "data/selection/full-AF.tsv", row.names = FALSE, 
            quote = FALSE, sep = "\t")

dir.create("data/selection/AF", showWarnings = FALSE)

for (sampleID in unique(mut_list$sampleID)) {
  sub_mut_list = mut_list[mut_list$sampleID == sampleID,]
  sub_mut_list = sub_mut_list[sub_mut_list$af >= 0.03,]
  write.table(sub_mut_list, paste0("data/selection/AF/AF-", sampleID, ".tsv"),
              row.names = FALSE, quote = FALSE, sep = "\t")
}
```

## Fit selection models

Please open the Mathematica notebook (`code/selection/fit-dist.nb`) and run it 
by hand (the outputs generated are used in subsequent cells). The notebook fits
the negative binomial model for neutral evolution.

In case you do not have access to Mathematica to run the notebook, we provide 
its output files in `data/selection/neg-bin-params-fit.csv` and 
`data/selection/neg-bin-rsquared-fit.csv`.

The code below runs the neutralitytestr model.

```{r}
fmin = 0.05
fmax = 0.45

petrAF = read.table("data/selection/full-AF.tsv", sep = "\t", header = T)
donors = unique(as.vector(petrAF$sampleID))

getSampleNtrtest <- function(afDF, sampleID, fmin, fmax){
  # run neutralitytestr on a single sample
  VAFsample = afDF[afDF$sampleID == sampleID, "af"]
  out = neutralitytest(VAFsample, fmin = fmin, fmax = fmax)
  results = c(sampleID,
              out$area$metric, out$area$pval,
              out$Dk$metric, out$Dk$pval,
              out$meanDist$metric, out$meanDist$pval,
              out$rsq$metric, out$rsq$pval,
              out$mutation.rate)
  names(results) = c("sampleID",
                     "area", "pval_area",
                     "Dk", "pval_Dk",
                     "meanDist", "pval_meanDist",
                     "rsq", "pval_rsq",
                     "mutrate")
  return(results)
}

ntrtestrPetrout = t(sapply(donors, 
                           function(sampleID) getSampleNtrtest(
                             petrAF, sampleID, fmin, fmax)))
write.table(ntrtestrPetrout, "data/selection/neutralitytestr.tsv", 
            sep = "\t", quote = FALSE, row.names = FALSE)

```

## Plot selection classification

Plot the selection classification based on the goodness of fit results from neutrality testr and the negative binomial like fit. 

```{r plot-prev, fig.height = 5, fig.width=7, warning=FALSE}
ntrtestrPetr = read.table("data/selection/neutralitytestr.tsv", 
                          stringsAsFactors = FALSE, header = TRUE)

negbinfitPetr = read.table("data/selection/neg-bin-rsquared-fit.csv",
                           stringsAsFactors = FALSE, header = TRUE, sep = ",")
negbinfitPetr$sampleID = negbinfitPetr$fname
negbinfitPetr$sampleID = gsub("AF-", "", negbinfitPetr$sampleID)
negbinfitPetr$sampleID = gsub(".tsv", "", negbinfitPetr$sampleID)
rownames(negbinfitPetr) = negbinfitPetr$sampleID

dfrsq = data.frame(sampleID = ntrtestrPetr$sampleID,
                   rsq_ntrtestr = ntrtestrPetr$rsq,
                   rsq_negbinfit = negbinfitPetr[ntrtestrPetr$sampleID, "rsq"])

cutoff_selection_cummut = 0.85
cutoff_selection_negbin = 0.25
cutoff_neutral_cummut = 0.9
cutoff_neutral_negbin = 0.55


donors = c("euts", "fawm", "feec", "fikt", "garx", "gesg", "heja", "hipn", "ieki",
           "joxm", "kuco", "laey", "lexy", "naju", "nusw", "oaaz", "oilg", "pipw",
           "puie", "qayj", "qolg", "qonc", "rozh", "sehl", "ualf", "vass", "vils",
           "vuna", "wahn", "wetu", "xugn", "zoxy")

dfrsq = dfrsq[(dfrsq$sampleID %in% donors),]

dfrsq$candidatelabel = NA
dfrsq$candidatelabel[dfrsq$sampleID == "puie"] = "puie"

filter_selection = (dfrsq$rsq_ntrtestr < cutoff_selection_cummut) & (dfrsq$rsq_negbinfit < cutoff_selection_negbin)
filter_neutral = (dfrsq$rsq_ntrtestr > cutoff_neutral_cummut) & (dfrsq$rsq_negbinfit > cutoff_neutral_negbin)

dfrsq$selection = "undetermined"
dfrsq$selection[filter_selection] = "selected"
dfrsq$selection[filter_neutral] = "neutral"


plt_scatter = ggplot(dfrsq, aes(x = rsq_negbinfit, y = rsq_ntrtestr)) +
  scale_colour_manual(values = c("neutral" = "#007536", "selected" = "#5EF288",
                               "undetermined" = "#CCCCCC")) +
  geom_point(aes(colour = selection)) +
  geom_label_repel(aes(label = candidatelabel), color = "white",
                   size = 2.5,
                   fill = "black", box.padding = 0.35, point.padding = 0.5,
                   segment.color = 'grey50') +
  theme_bw() +
  theme(text = element_text(size = 9), axis.text = element_text(size = 8),
        axis.title = element_text(size = 9), 
        plot.title = element_text(size = 9, hjust = 0.5)) +
  labs(x = "Goodness of Fit - Negative Binomial Distribution", 
       y = "Goodness of Fit - Cumulative Mutations") +
  theme(strip.background = element_blank()) +
  labs(title = "") +
  theme(legend.justification = c(1,0), legend.position = c(1,0)) +
  theme(legend.background = element_rect(fill = "transparent", 
                                         colour = "transparent"), 
        legend.key.size = unit(0.25, "cm")) +
  labs(colour = "Selection")

plt_scatter
```


## Visualise negative binomial model

Results from the negative binomial model for neutral evolution.

```{r plot-neg-bin-fit, fig.height = 20, fig.width=17}
dfAFpetr = read.table("data/selection/full-AF.tsv", sep = "\t", 
                      stringsAsFactors = FALSE, header = TRUE)
dfAFpetr = dfAFpetr[dfAFpetr$sampleID %in% donors,]

fitparamspetr = read.csv("data/selection/neg-bin-params-fit.csv", 
                         stringsAsFactors = FALSE, header = TRUE)
fitparamspetr$sampleID = gsub("AF-", "", fitparamspetr$fname)
fitparamspetr$sampleID = gsub(".tsv", "", fitparamspetr$sampleID)

a = 1
b = 1
fun.1 <- function(x, a=1, b=1) 1/(a*x)*exp(-x/b)
args = list(mean = 2, sd = .5)

dd <- data.frame(
  predicted = rnorm(72, mean = 2, sd = 2),
  state = rep(c("A", "B", "C"), each = 24)
)

# save generate plotting points
grid = seq(0.01, 0.6, length = 500)

normaldens <- ddply(fitparamspetr, "sampleID", function(df) {
  data.frame(
    predicted = grid,
    density = fun.1(grid, df$a, df$b)
  )
})

normaldens = normaldens[normaldens$sampleID %in% donors,]

dfAFpetr$id = sapply(dfAFpetr$sampleID, 
                     function(sampleID) paste0(
                       sampleID, " (rsq ",
                       round(fitparamspetr$rsq[fitparamspetr$sampleID ==
                                                 sampleID],2),")"))
normaldens$id = sapply(normaldens$sampleID, 
                       function(sampleID) dfAFpetr$id[dfAFpetr$sampleID ==
                                                        sampleID][1])

plt_hist = ggplot(dfAFpetr, aes(x = af)) +
  geom_vline(xintercept = fmin, colour = "grey") +
  geom_vline(xintercept = fmax, colour = "grey") +
  geom_histogram(binwidth = (fmax - fmin)/40) +
  geom_line(aes(x = predicted, y = density), data = normaldens, colour = "red") +
  facet_wrap(~ id, scales = "free_y", ncol = 4) +
  theme_bw() +
  theme(text = element_text(size = 9), axis.text = element_text(size = 8), axis.title = element_text(size = 9), plot.title = element_text(size = 9, hjust = 0.5)) +
  labs(x = paste0("Allele Frequency"), y = "# Mutations") +
  theme(legend.position = "none") +
  labs(colour = "") +
  theme(strip.background = element_blank()) +
  labs(title = "")


pname = paste0("selection-neg-bin-fit")
ppath = paste0("figures/selection/", pname, ".pdf")
ggsave(ppath, plot = plt_hist, width = 17, height = 20, units  =  "cm")
ppath  =  paste0("figures/selection/", pname, ".png")
ggsave(ppath, plot = plt_hist, width = 17, height = 20, dpi = 300, units = "cm", limitsize = FALSE)
plt_hist
```


## Visualise results from neutralitytestr

Results from the cumulative mutations model for neutral evolution.

```{r plot-neutralitytestr, fig.height = 25, fig.width=17}
dfntrtestr = read.table("data/selection/neutralitytestr.tsv", 
                        stringsAsFactors = FALSE, header = TRUE)

plotSampleCumMut <- function(afDF, dfntrtestr, sampleID, fmin, fmax) {
  # plot cummulative mutations as per Sottoriva & Graham
  afDFsample = afDF[afDF$sampleID == sampleID, ]
  rsq = dfntrtestr$rsq[dfntrtestr$sampleID == sampleID]
  # cumsum with decreasing frequency
  afDFsample = afDFsample[order(afDFsample$af, decreasing = TRUE), ]
  afDFsample$cumsum = 1:length(afDFsample$af)
  afDFsample$inverse_af = 1/afDFsample$af
  plt_cummut = ggplot(afDFsample, aes(x = inverse_af, y = cumsum)) +
    geom_vline(xintercept = c(1/fmin, 1/fmax), colour = "darkgrey") +
    geom_point(size = 0.5) +
    geom_line(data = subset(afDFsample, 
                            (inverse_af < 1/fmin) & (inverse_af > 1/fmax)),
              stat = "smooth", method = 'lm', formula = y ~ x, se = FALSE, 
              colour = "red", alpha = 0.5, size = 0.8) +
    coord_cartesian(xlim = c(0, 1/0.01)) +
    theme_bw() +
    theme(text = element_text(size = 9), axis.text = element_text(size = 8),
          axis.title = element_text(size = 9), 
          plot.title = element_text(size = 9, hjust = 0.5)) +
    labs(x = paste0("Inverse AF"), y = "# Cum Mut") +
    theme(legend.position = "none") +
    # remove unnecessary facet
    theme(strip.background = element_blank()) +
    labs(title = paste0(sampleID, " (rsq ", round(rsq,2), ")"))
  return(plt_cummut)
}


pltsCumMut = cowplot::plot_grid(
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[1], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[2], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[3], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[4], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[5], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[6], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[7], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[8], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[9], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[10], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[11], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[12], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[13], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[14], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[15], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[16], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[17], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[18], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[19], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[20], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[21], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[22], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[23], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[24], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[25], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[26], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[27], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[28], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[29], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[30], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[31], fmin, fmax),
  plotSampleCumMut(dfAFpetr, dfntrtestr, donors[32], fmin, fmax),
  ncol = 4)

pname = paste0("selection-neutralitytestr")
ppath = paste0("figures/selection/", pname, ".pdf")
ggsave(ppath, plot = pltsCumMut, width = 17, height = 25, units = "cm")
ppath = paste0("figures/selection/", pname, ".png")
ggsave(ppath, plot = pltsCumMut, width = 17, height = 25, dpi = 300, 
       units = "cm", limitsize = FALSE)
pltsCumMut
```


