---
title: "Selection Models"
author: "Daniel J. Kunz & Davis J. McCarthy"
site: workflowr::wflow_site
---

 ## Load libraries and data
```{r setup, include=TRUE, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
library(cowplot)
library(plyr)
dir.create("figures/selection", showWarnings = FALSE, recursive = TRUE)
```

Load the call set and extract the allele frequencies which are used for the fits of the selection models.
```{r}
filteredAF = read.table("data/exome-point-mutations/high-vs-low-exomes.v62.ft.filt_lenient-alldonors.txt.gz", 
                        header = TRUE, stringsAsFactors = FALSE)
 mut_list = data.frame("sampleID" = filteredAF$donor_short_id,
                      "af_fibro" = filteredAF$nALT_fibro/(filteredAF$nREF_fibro + filteredAF$nALT_fibro),
                      "af_ips" = filteredAF$nALT_ips/(filteredAF$nREF_ips + filteredAF$nALT_ips),
                      "chr" = filteredAF$chrom,
                      "pos" = filteredAF$pos,
                      "ref" = filteredAF$ref,
                      "mut" = filteredAF$alt,
                      "mutID" = paste(filteredAF$chrom, filteredAF$pos, filteredAF$ref, filteredAF$alt, sep = "_"))
mut_list = mut_list[order(mut_list$sampleID),]
write.table(mut_list, "data/selection/ips-fibro-AF.tsv", 
            row.names = FALSE, quote = FALSE, sep = "\t")
 mut_list = data.frame("sampleID" = filteredAF$donor_short_id,
                      "af" = filteredAF$nALT_fibro/(filteredAF$nREF_fibro + filteredAF$nALT_fibro),
                      "chr" = filteredAF$chrom,
                      "pos" = filteredAF$pos,
                      "ref" = filteredAF$ref,
                      "mut" = filteredAF$alt)
mut_list = mut_list[order(mut_list$sampleID),]
 write.table(mut_list, "data/selection/full-AF.tsv", row.names = FALSE, 
            quote = FALSE, sep = "\t")
 dir.create("data/selection/AF", showWarnings = FALSE)
 for (sampleID in unique(mut_list$sampleID)) {
  sub_mut_list = mut_list[mut_list$sampleID == sampleID,]
  sub_mut_list = sub_mut_list[sub_mut_list$af >= 0.03,]
  write.table(sub_mut_list, paste0("data/selection/AF/AF-", sampleID, ".tsv"),
              row.names = FALSE, quote = FALSE, sep = "\t")
}
```
 ## Fit selection models

For the selection analysis SubConalSelection (http://dx.doi.org/10.1038/s41588-018-0128-6) was used.
To reproduce the analysis please run the Julia code (`code/selection/subclonal-bayesian-ABC.jl`). 

Since the simulations take in the order of days/weeks we provide the outputfiles of the simulation in `data/subclonal-output-1/` and a summary in `data/p1-selection.csv`.

 ## Plot selection classification
 Plot the selection classification from SubConalSelection. The grey background indicates results with high uncertainty due to low numbers of mutations (< 100).
 
```{r plot-selection-results, fig.height = 5, fig.width=7, warning=FALSE}
donors = c("joxm", "garx", "wahn", "vass", "ualf", "euts", "laey", "pipw", "oilg", "heja",
          "sehl", "feec", "gesg", "fikt", "vuna", "qonc", "xugn", "qolg", "puie", "fawm",
          "oaaz", "naju", "ieki", "rozh", "wetu", "nusw", "zoxy", "hipn", "lexy", "vils",
          "qayj", "kuco")

dfResults = read.csv("data/p1-selection.csv", stringsAsFactors=F)

dfResults$donor = factor(dfResults$donor, levels=rev(donors[donors %in% dfResults$donor]))

plt_scatter = ggplot(dfResults, aes(x=ps1, y=donor)) +
                geom_point(alpha=0) + # workaround to plot background
                annotate("rect", xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=12.5, alpha=0.2, fill="grey") +
                geom_point(aes(colour=selection)) +
                coord_cartesian(xlim = c(0, 1)) +
                scale_colour_manual(values = c(neutral="#1283FF", selected="#144E7B", undetermined="#CACACA")) +
                geom_vline(xintercept=c(0.3, 0.7), colour="#808080") +
                scale_x_continuous(breaks=c(0, 0.3, 0.5, 0.7, 1.0)) +
                theme_bw() +
                theme(text=element_text(size=7), axis.text=element_text(size=6), axis.title=element_text(size=7), plot.title=element_text(size=7, hjust=0.5)) +
                labs(x="Probability of Selection", y="") +
                theme(legend.position="none") +
                # remove unnecessary facet
                theme(strip.background = element_blank()) +
                theme(legend.position="top") +
                labs(title="") +
                labs(colour="Selection")

plt_scatter
```

 ## Plot fit of selection models
 Plot the fit of the models to the allele frequency distribution.
 
```{r plot-nodel-fit, fig.height = 20, fig.width=17}

selection_class = read.csv("data/p1-selection.csv", stringsAsFactors=F)

# histograms
library(ggplot2)

fout = list.files(path = "data/subclonal-output-1", pattern="*histogram-clone1[.]*",
                  full.names=T, recursive=T)
donors_neutral = selection_class[selection_class$selection == "neutral", "donor"]

for (donor in donors_neutral){
  fout[grepl(donor, fout)] = gsub("clone1", "clone0", fout[grepl(donor, fout)])
}


modelstats = read.csv(fout[1])
modelstats$donor = strsplit(basename(fout[1]), "-")[[1]][3]
modelstats$selection_model = gsub(".csv", "" , strsplit(basename(fout[1]), "-")[[1]][5])

for (i in 2:length(fout)){
  modelstats_donor = read.csv(fout[i])
  modelstats_donor$donor = strsplit(basename(fout[i]), "-")[[1]][3]
  modelstats_donor$selection_model = gsub(".csv", "" , strsplit(basename(fout[i]), "-")[[1]][5])
  modelstats = rbind(modelstats, modelstats_donor)
}

modelstats$selection_model = gsub("clone0", "neutral", modelstats$selection_model)
modelstats$selection_model = gsub("clone1", "selected", modelstats$selection_model)

selection_class$donor_class = paste0(selection_class$donor, " (", selection_class$selection, ")")

# add to modelstats
modelstats$donor_class = NA
modelstats$selection = NA

for (donor in selection_class$donor){
  modelstats[modelstats$donor == donor, "donor_class"] = selection_class[selection_class$donor == donor, "donor_class"]
  modelstats[modelstats$donor == donor, "selection"] = selection_class[selection_class$donor == donor, "selection"]
}


fmin = 0.05
fmax = 0.45


# remove model fit for undetermined donors
modelstats[modelstats$selection == "undetermined", c("mean", "lowerq95", "upperq95")] = NA

plt_hist = ggplot(modelstats, aes(x=VAF, y=truecounts)) +
                facet_wrap(~donor_class, ncol=4, scales = "free_y") +
                geom_bar(stat="identity") +
                geom_line(aes(x=VAF, y=mean, colour=selection_model)) +
                geom_ribbon(aes(x=VAF, ymax=upperq95, ymin=lowerq95, fill=selection_model), alpha=0.2) +
                geom_vline(xintercept=c(fmin, fmax), colour="#808080") +
                # scale_x_continuous(limits = c(0, 0.47), breaks=c(0, fmin, 0.1, 0.2, 0.3, 0.4, fmax), expand=c(0,0)) +
                coord_cartesian(xlim = c(0, 0.47), ylim=c(0, max(modelstats$truecounts)), expand=0) +
                scale_colour_manual(values = c(neutral="#1283FF", selected="#144E7B", undetermined="#CACACA"), guide=FALSE) +
                scale_fill_manual(values = c(neutral="#1283FF", selected="#144E7B", undetermined="#CACACA")) +
                scale_x_continuous(breaks=c(0, fmin, 0.1, 0.2, 0.3, 0.4, fmax), labels=c("0", paste0(fmin), "0.1", "0.2", "0.3", "0.4", paste0(fmax))) +
                theme_bw() +
                theme(text=element_text(size=7), axis.text=element_text(size=6), axis.title=element_text(size=7), plot.title=element_text(size=7, hjust=0.5)) +
                labs(x="VAF", y="# Mutations") +
                # remove unnecessary facet
                theme(strip.background = element_blank()) +
                theme(legend.position="bottom") +
                labs(title="") +
                labs(fill="Selection Model")
                # coord_fixed()

ppath = paste0("figures/selection-hist-SubClonalSelection.png")
ggsave(ppath, plot=plt_hist, width=18.3, height=20, dpi=300, units = "cm")
ppath = paste0("figures/selection-hist-SubClonalSelection.pdf")
ggsave(ppath, plot=plt_hist, width=18.3, height=20, units = "cm")
plt_hist
```
