---
title: "Simulation analysis: systematic and for all lines"
author: "Yuanhua Huang & Davis J. McCarthy"
site: workflowr::wflow_site
---

## Load libraries and simulation results

```{r load-libs, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
dir.create("figures/simulations", showWarnings = FALSE, recursive = TRUE)
library(ggpubr)
library(tidyverse)
library(cardelino)
library(viridis)
library(cowplot)
library(latex2exp)
library(ggrepel)
lines <- c("euts", "fawm", "feec", "fikt", "garx", "gesg", "heja", "hipn", 
            "ieki", "joxm", "kuco", "laey", "lexy", "naju", "nusw", "oaaz", 
            "oilg", "pipw", "puie", "qayj", "qolg", "qonc", "rozh", "sehl", 
            "ualf", "vass", "vuna", "wahn", "wetu", "xugn", "zoxy", "vils")
```


## Simulation experiments

Define functions for simulation.

```{r}
# assess cardelino with simulation
dat_dir <- "data/"
data(config_all)
data(simulation_input)
simu_input <- list("D" = D_input)

get_prob_label <- function(X){
  max_idx <- rep(0, nrow(X))
  for (j in seq_len(nrow(X))) {
    max_idx[j] <- which.max(X[j, ])
  }
  max_idx
}

get_prob_value <- function(X, mode = "best") {
  max_val <- rep(0, nrow(X))
  for (i in seq_len(nrow(X))) {
    sorted_val <- sort(X[i,], decreasing = TRUE)
    if (mode == "delta") {
      max_val[i] <- sorted_val[1] - sorted_val[2]
    } else if (mode == "second") {
      max_val[i] <- sorted_val[2]
    } else {#default mode: best
      max_val[i] <- sorted_val[1]
    }
  }
  max_val
}

demuxlet <- function(A, D, Config, theta1 = 0.5, theta0 = 0.01) {
  P0_mat <- dbinom(A, D, theta0, log = TRUE)
  P1_mat <- dbinom(A, D, theta1, log = TRUE)
  P0_mat[which(is.na(P0_mat))] <- 0
  P1_mat[which(is.na(P1_mat))] <- 0
  logLik_mat <- t(P0_mat) %*% (1 - Config) + t(P1_mat) %*% Config
  prob_mat <- exp(logLik_mat) / rowSums(exp(logLik_mat))
  prob_mat
}

simulate_joint <- function(Config_all, D_all, n_clone = 4, mut_size = 5,
                           missing = NULL, error_mean = c(0.01, 0.44), 
                           error_var = c(30, 4.8), n_repeat = 1) {
  simu_data_full <- list()
  for (i in seq_len(n_repeat)) {
    Config <- sample(Config_all[[n_clone - 2]], size = 1)[[1]]
    Config <- matrix(rep(c(t(Config)), mut_size), ncol = ncol(Config), 
                     byrow = TRUE)
    row.names(Config) <- paste0("SNV", seq_len(nrow(Config)))
    colnames(Config) <- paste0("Clone", seq_len(ncol(Config)))
    D_input <- sample_seq_depth(D_all, n_cells = 200, missing_rate = missing, 
                                n_sites = nrow(Config))
    sim_dat <- sim_read_count(Config, D_input, Psi = NULL, cell_num = 200, 
                              means = error_mean, vars = error_var)
    sim_dat[["Config"]] <- Config
    simu_data_full[[i]] <- sim_dat
  }
  simu_data_full
}

assign_score <- function(prob_mat, simu_mat, threshold=0.2, mode="delta") {
  assign_0 <- get_prob_label(simu_mat)
  assign_1 <- get_prob_label(prob_mat)
  prob_val <- get_prob_value(prob_mat, mode = mode)
  idx <- prob_val >= threshold
  
  rt_list <- list("ass" = mean(idx), 
                  "acc" = mean(assign_0 == assign_1),
                  "acc_ass" = mean((assign_0 == assign_1)[idx]))
  rt_list
}

assign_curve <- function(prob_mat, simu_mat, mode="delta"){
  assign_0 <- get_prob_label(simu_mat)
  assign_1 <- get_prob_label(prob_mat)
  prob_val <- get_prob_value(prob_mat, mode = mode)
  
  thresholds <- sort(unique(prob_val))
  ACC <- rep(0, length(thresholds))
  ASS <- rep(0, length(thresholds))
  for (i in seq_len(length(thresholds))) {
    idx <- prob_val >= thresholds[i]
    ASS[i] <- mean(idx)
    ACC[i] <- mean((assign_0 == assign_1)[idx])
  }
  thresholds <- c(thresholds, 1.0)
  ACC <- c(ACC, 1.0)
  ASS <- c(ASS, 0.0)
  AUC <- AUC_acc_ass <- 0.0
  for (i in seq_len(length(thresholds) - 1)) {
    AUC <- AUC + 0.5 * (thresholds[i] - thresholds[i + 1]) * 
      (ACC[i] + ACC[i + 1])
    AUC_acc_ass <- AUC_acc_ass + 0.5 * (ASS[i] - ASS[i + 1]) * 
      (ACC[i] + ACC[i + 1])
  }
  AUC <- AUC / (thresholds[1] - thresholds[length(thresholds)])
  AUC_acc_ass <- AUC_acc_ass / (ASS[1] - ASS[length(thresholds)])
  
  rt_list <- list("ACC" = ACC, "ASS" = ASS, "AUC" = AUC, 
                  "AUC_acc_ass" = AUC_acc_ass, "thresholds" = thresholds)
  rt_list
}

assign_macro_ROC <- function(prob_mat, simu_mat) {
  thresholds <- seq(0, 0.999, 0.001)
  ACC <- rep(0, length(thresholds))
  ASS <- rep(0, length(thresholds))
  FPR <- rep(0, length(thresholds))
  TPR <- rep(0, length(thresholds))
  for (i in seq_len(length(thresholds))) {
    idx <- prob_mat >= thresholds[i]
    ASS[i] <- mean(idx) # not very meaningful
    ACC[i] <- mean(simu_mat[idx])
    FPR[i] <- sum(simu_mat[idx] == 0) / sum(simu_mat == 0)
    TPR[i] <- sum(simu_mat[idx] == 1) / sum(simu_mat == 1)
  }
  
  AUC <- 0.0
  for (i in seq_len(length(thresholds) - 1)) {
    AUC <- AUC + 0.5 * (FPR[i] - FPR[i + 1]) * (TPR[i] + TPR[i + 1])
  }
  
  rt_list <- list("FPR" = FPR, "TPR" = TPR, "AUC" = AUC,
                  "thresholds" = thresholds, "ACC" = ACC, "ASS" = ASS)
  rt_list
}

```

Run simulations.

```{r run-sims}
set.seed(1)
ACC_all <- c()
ASS_all <- c()
AUC_all <- c()
ERR_all <- c()
labels_all <- c()
method_all <- c()
variable_all <- c()

type_use <- c("mut_size", "n_clone", "missing", "FNR", "shapes1")
value_list <- list(c(3, 5, 7, 10, 15, 25), 
                   seq(3, 8), 
                   seq(0.7, 0.95, 0.05), 
                   seq(0.35, 0.6, 0.05),
                   c(0.5, 1.0, 2.0, 4.0, 8.0, 16.0))

for (mm in seq_len(length(value_list))) {
  values_all <- value_list[[mm]]
  for (k in seq_len(length(values_all)))  {
    if (mm == 1) {
      simu_data <- simulate_joint(Config_all, simu_input$D, n_clone = 4, 
                                  mut_size = values_all[k], missing = 0.8, 
                                  error_mean = c(0.01, 0.44), n_repeat = 20)
    } else if (mm == 2) { 
      simu_data <- simulate_joint(Config_all, simu_input$D, 
                                  n_clone = values_all[k], mut_size = 10, 
                                  missing = 0.8, error_mean = c(0.01, 0.44),
                                  n_repeat = 20)
    } else if (mm == 3) {
      simu_data <- simulate_joint(Config_all, simu_input$D, n_clone = 4, 
                                  mut_size = 10, missing = values_all[k], 
                                  error_mean = c(0.01, 0.44), n_repeat = 20)
    } else if (mm == 4) {
      simu_data <- simulate_joint(Config_all, simu_input$D, n_clone = 4, 
                                  mut_size = 10, missing = 0.8, 
                                  error_mean = c(0.01, values_all[k]), 
                                  n_repeat = 20)
    } else if (mm == 5) {
      simu_data <- simulate_joint(Config_all, simu_input$D, n_clone = 4, 
                                  mut_size = 10, missing = 0.8, 
                                  error_mean = c(0.01, 0.44), 
                                  error_var = c(30, values_all[k]), 
                                  n_repeat = 20)
    }
    
    for (d_tmp in simu_data) {
      prob_all <- list()
      methods_use <- c("demuxlet", "Bern_EM", "Binom_EM", "Binom_Gibbs",
                       "Binom_gmline")
      prob_all[[1]] <- demuxlet(d_tmp$A_sim, d_tmp$D_sim, d_tmp$Config, 
                                theta0 = mean(d_tmp$theta0_binom, na.rm = TRUE))
      prob_all[[2]] <- cell_assign_EM(d_tmp$A_sim, d_tmp$D_sim, d_tmp$Config, 
                                      Psi = NULL, verbose = F)$prob
      prob_all[[3]] <- cell_assign_EM(d_tmp$A_sim, d_tmp$D_sim, d_tmp$Config, 
                                      Psi = NULL, model = "binomial",
                                      verbose = FALSE)$prob
      prob_all[[4]] <- cell_assign_Gibbs(d_tmp$A_sim, d_tmp$D_sim,
                                         d_tmp$Config, Psi = NULL,
                                         min_iter = 1000, wise = "variant", 
                                         prior1 = c(2.11, 2.69), 
                                         verbose = FALSE)$prob
      
      for (i in seq_len(length(prob_all))) {
        prob_mat <- prob_all[[i]]
        assign_scr <- assign_score(prob_mat, d_tmp$I_sim, 
                                                threshold = 0.5001, 
                                                mode = "best")
        ACC_all <- c(ACC_all, assign_scr$acc_ass)
        ASS_all <- c(ASS_all, assign_scr$ass)
        AUC_all <- c(AUC_all, assign_curve(prob_mat, d_tmp$I_sim, 
                                           mode = "best")$AUC_acc_ass)
        ERR_all <- c(ERR_all, mean(abs(prob_mat - d_tmp$I_sim)))
        labels_all <- c(labels_all, values_all[k])
        method_all <- c(method_all, methods_use[i])
        variable_all <- c(variable_all, type_use[mm])
      }
    }
  }
}

df <- data.frame(Accuracy = ACC_all, AUC_of_ACC_ASS = AUC_all, 
                 Assignable = ASS_all, MAE = ERR_all, Methods = method_all, 
                 labels = labels_all, variable = variable_all)

dat_dir <- "data/simulations"
saveRDS(df, paste0(dat_dir, "/simulate_extra_s1_v2.rds"))
```


## Precision-recall curve

```{r pr-curve-calc}
set.seed(1)
PCAU_curve_list <- list()

simu_data <- simulate_joint(Config_all, simu_input$D, n_clone = 4, 
                              mut_size = 10, missing = 0.8, 
                              error_mean = c(0.01, 0.44), n_repeat = 20)

assign_0 <- matrix(0, nrow = 200, ncol = 20)
assign_1 <- matrix(0, nrow = 200, ncol = 20)
prob_all <- matrix(0, nrow = 200, ncol = 20)
for (i in seq_len(length(simu_data))) {
  d_tmp <- simu_data[[i]]
  prob_tmp <- cell_assign_Gibbs(d_tmp$A_sim, d_tmp$D_sim,
                                d_tmp$Config, Psi = NULL,
                                min_iter = 1000, wise = "variant", 
                                prior1 = c(2.11, 2.69), verbose = FALSE)$prob
  assign_0[, i] <- get_prob_label(d_tmp$I_sim)
  assign_1[, i] <- get_prob_label(prob_tmp)
  prob_all[, i] <- get_prob_value(prob_tmp, mode = "best")
}

dat_dir = "data/simulations"
saveRDS(list("assign_0" = assign_0, "assign_1" = assign_1, 
             "prob_all" = prob_all), 
        paste0(dat_dir, "/simulate_prob_curve.rds"))
```


## Analysis of systematic results

### Precision-recall curve for default settings

```{r}
dat_dir = "data/simulations"
rds.tmp <- readRDS(paste0(dat_dir, "/simulate_prob_curve.rds"))
assign_0 <- rds.tmp$assign_0
assign_1 <- rds.tmp$assign_1
prob_all <- rds.tmp$prob_all

thresholds <- seq(0, 1, 0.001)
recalls <- rep(0, length(thresholds))
precision_all <- matrix(0, nrow = length(thresholds), ncol = ncol(prob_all) + 1)
for (i in seq_len(length(thresholds))) {
    idx <- prob_all >= thresholds[i] 
    recalls[i] <- mean(idx)
    precision_all[i, ncol(prob_all) + 1] <- mean((assign_0 == assign_1)[idx])
    for (j in seq_len(ncol(prob_all))) {
        idx <- prob_all[, j] >= sort(prob_all[, j], 
                                     decreasing = TRUE)[round(recalls[i] *
                                                                nrow(prob_all))]
        precision_all[i, j] <- mean((assign_0[,j] == assign_1[,j])[idx])
    }
}

order_idx <- order(colMeans(precision_all[, 1:ncol(prob_all)]))
idx1 <- order_idx[round(0.25 * length(order_idx))]
idx2 <- order_idx[round(0.75 * length(order_idx))]

df.tmp <- data.frame(cutoff = thresholds, Recall = recalls, 
                     Presision = precision_all[, ncol(precision_all)], 
                     ACC_low1 = precision_all[, idx1],
                     ACC_high1 = precision_all[, idx2])
```

Calculate AUC score.

```{r}
nn <- ncol(precision_all)
AUC_score <- 0.0
for (i in seq_len(length(recalls) - 1)) {
    AUC_score <- AUC_score + 0.5 * (recalls[i] - recalls[i + 1]) * 
      (precision_all[i, nn] + precision_all[i + 1, nn])
}
AUC_score <- AUC_score / (recalls[1] - recalls[length(recalls)])
print(AUC_score)
```

Plot PR curve.

```{r plot-pr-curve, fig.height=2.5, fig.width=3.5}
fig_dir <- "figures/simulations"

idx_05 <- prob_all >= 0.5
recall_05 <- mean(idx_05)
precision_05 <- mean((assign_0 == assign_1)[idx_05])
print(c(recall_05, precision_05))

fig.curve <- ggplot(df.tmp, aes(x = Recall, y = Presision)) + 
  geom_ribbon(aes(ymin = ACC_low1, ymax = ACC_high1), fill = "grey80") +
  scale_color_viridis(option = "B") +
  geom_line(aes(color = cutoff)) + geom_point(aes(color = cutoff), size = 0.5) + 
  geom_point(aes(x = recall_05, y = precision_05), shape = 1, color = "black",
             size = 3) +
  xlab("Recall: Fraction of assignable cells") + 
  ylab("Precision") + 
  ylim(0.5, 1) +
  pub.theme() + 
  theme(legend.position = c(0.25,0.45)) + 
  labs(color = 'highest P') 

ggsave(paste0(fig_dir, "/fig1b_PRcurve.png"), 
       fig.curve, height = 2.5, width = 3.5, dpi = 300)
ggsave(paste0(fig_dir, "/fig1b_PRcurve.pdf"), 
       fig.curve, height = 2.5, width = 3.5, dpi = 300)
fig.curve
```


### Comparison of parameters

Assess cardelino with simulated data.

```{r}
df <- readRDS(paste0("data/simulations/simulate_extra_s1_v2.rds"))

## Change method names
df$Methods <- as.character(df$Methods)
df <- df[df$Methods != "Bern_EM", ]
df$Methods[df$Methods == "demuxlet"] <- "theta_fixed"
df$Methods[df$Methods == "Binom_EM"] <- "theta_EM"
df$Methods[df$Methods == "Binom_Gibbs"] <- "cardelino"
df$Methods <- as.factor(df$Methods)

## Change variables
df$labels[df$variable == "missing"] <- 1 - df$labels[df$variable == "missing"]
df$labels[df$variable == "shapes1"] <- 1 / df$labels[df$variable == "shapes1"] #paste0("1/", df$labels[df$variable == "shapes1"])
```

Table of results from simulations:

```{r}
head(df)
```

Area under Precision-Recall curves for different parameter settings.

```{r fig1d, fig.height=2.5, fig.width=3.5}
fig_dir <- "figures/simulations/"

df1 <- df[df$Methods == "cardelino", ]

type_use <- c("mut_size", "n_clone", "missing", "FNR", "shapes1")
xlabels <- c("# variants per clonal branch", "Number of clones", 
             "Overall variant coverage")
titles <- c("Mutations per branch", "Numbers of clones", "Variant coverages",
            "Fraction of ALT allele", "Constrentration of allelic expr")

df1 <- df1[df1$labels != 40, ]
df1 <- df1[df1$labels != 0.65, ]
df1 <- df1[df1$labels != 0.60, ]

fig_list <- list()
for (mm in c(1,3)) {
  fig_list[[mm]] <- ggplot(df1[df1$variable == type_use[mm], ],
                           aes(x = as.factor(labels), y = AUC_of_ACC_ASS)) +
    ylab("AU precision-recall curve") + 
    geom_boxplot() + xlab(xlabels[mm]) + 
    ylim(0.6, 1) + pub.theme()
  #     if (mm == 1) {
  #         fig_list[[mm]] <- fig_list[[mm]] + 
  #             scale_x_discrete(labels=c("3", "5", "7", "10*", "15", "15")) }
  #     if (mm == 3) {
  #         fig_list[[mm]] <- fig_list[[mm]] + 
  #             scale_x_discrete(labels=c("0.05", "0.1", "0.15", "0.2*", "0.25", "0.3")) }
}

ggsave(file = paste0(fig_dir, "fig_1c_mutations.png"),
       fig_list[[1]], height = 2.5, width = 3.5, dpi = 300)
ggsave(file = paste0(fig_dir, "fig_1c_mutations.pdf"),
       fig_list[[1]], height = 2.5, width = 3.5, dpi = 300)

ggsave(file = paste0(fig_dir, "fig_1d_coverages.png"),
       fig_list[[3]], height = 2.5, width = 3.5, dpi = 300)
ggsave(file = paste0(fig_dir, "fig_1d_coverages.pdf"),
       fig_list[[3]], height = 2.5, width = 3.5, dpi = 300)

fig_list[[1]]
fig_list[[3]]
```

Boxplots comparing different models for cell-clone assignment.

```{r, fig.height=7, fig.width=12}
type_use <- c("mut_size", "n_clone", "missing", "FNR", "shapes1")
xlabels <- c("# variants per clonal branch", "Number of clones", 
             "Overall variant coverage", 
             "Mean fraction of ALT alleles", "Variance of allelic imbalance")
df$Methods <- ordered(df$Methods, levels=c("theta_fixed", "theta_EM", 
                                           "cardelino")) #Bern_EM

xlabel_list <- list(c("3", "5", "7", "10*", "15", "15"), 
                    c("3", "4*", "5", "6", "7", "8"),
                    c("0.05", "0.1", "0.15", "0.2*", "0.25", "0.3"),
                    c("0.35", "0.4", "* 0.45", "0.5", "0.55", "0.6"),
                    c("1/16", "1/8", "*  1/4", "1/2", "1", "2"))

fig_list <- list()
for (mm in seq_len(length(type_use))) {
  fig_list[[mm]] <- ggplot(df[df$variable == type_use[mm], ],
                           aes(x = as.factor(labels), y = AUC_of_ACC_ASS,
                               fill=Methods)) + 
    geom_boxplot() + xlab(xlabels[mm]) + ylab("AUC: precision-recall curve") +
    pub.theme() + 
    scale_fill_brewer() + 
    scale_x_discrete(labels=xlabel_list[[mm]])
}

fig_box <- ggarrange(fig_list[[1]], fig_list[[3]], fig_list[[2]], 
                     fig_list[[4]], fig_list[[5]], 
                     labels = c("A", "B", "C", "D", "E"),
                     nrow = 2, ncol = 3, align = "hv", 
                     common.legend = TRUE, legend = "bottom")

ggsave(file = paste0(fig_dir, "simulation_overall_AUC.png"),
       fig_box, height = 7, width = 12, dpi = 300)
ggsave(file = paste0(fig_dir, "simulation_overall_AUC.pdf"),
       fig_box, height = 7, width = 12, dpi = 300)

fig_box
```


## Analysis of simulation results for individual lines

### Load simulation results for individual lines

```{r}
all_files <- paste0(lines, ".simulate.rds")
assign_0 <- matrix(0, nrow = 500, ncol = length(lines))
assign_1 <- matrix(0, nrow = 500, ncol = length(lines))
prob_all <- matrix(0, nrow = 500, ncol = length(lines))

for (i in seq_len(length(all_files))) {
  afile <- all_files[i]
  sim_dat <- readRDS(file.path("data", "simulations", afile))
  assign_0[, i] <- get_prob_label(sim_dat$I_sim)
  assign_1[, i] <- get_prob_label(sim_dat$prob_Gibbs)
  prob_all[, i] <- get_prob_value(sim_dat$prob_Gibbs, mode = "best")
}
```


### Load results from real data

#### Cardelino-original

```{r}
all_files <- paste0("cardelino_results.", lines, 
                    ".filt_lenient.cell_coverage_sites.rds")
n_sites <- rep(0, length(lines))
n_clone <- rep(0, length(lines))
recall_all <- rep(0, length(lines))
for (i in seq_len(length(all_files))) {
  afile <- all_files[i]
  carde_dat <- readRDS(file.path("data", "cell_assignment", afile))
  n_sites[i] <- nrow(carde_dat$D)
  n_clone[i] <- ncol(carde_dat$prob_mat)
  recall_all[i] <- mean(get_prob_value(carde_dat$prob_mat, mode = "best") > 0.5)
}
```


#### Cardelino-relax

```{r}
all_files_carderelax <- paste0("cardelino_results_carderelax.", lines, 
                    ".filt_lenient.cell_coverage_sites.rds")
recall_all_carderelax <- rep(0, length(lines))
for (i in seq_len(length(all_files))) {
  afile <- all_files_carderelax[i]
  carde_dat <- readRDS(file.path("data", "cell_assignment", afile))
  recall_all_carderelax[i] <- mean(get_prob_value(carde_dat$prob_mat, mode = "best") > 0.5)
}
```


#### Combine results

Overall correlation in assignment rates (recall) from simulated and observed 
data is `r round(cor(colMeans(prob_all > 0.5), recall_all), digits = 3)`.


```{r}
precision_simu <- rep(0, length(lines))
for (i in seq_len(length(lines))) {
  idx <- prob_all[, i] > 0.5
  precision_simu[i] <- mean(assign_0[idx, i] == assign_1[idx, i])
}

df <- data.frame(line = lines, n_sites = n_sites, n_clone = n_clone, 
                 recall_real_cardeorig = recall_all, 
                 recall_real_carderelax = recall_all_carderelax,
                 recall_simu = colMeans(prob_all > 0.5),
                 precision_simu = precision_simu)
```


## Plot observed vs simulated assignment rates (recall)

### Cardelino-original

```{r}
p <- df %>%
  dplyr::mutate(labs = ifelse(line == "joxm", "joxm", "")) %>%
  dplyr::mutate(sites_per_clone = cut(n_sites / pmax(n_clone - 1, 1), 
                                      breaks = c(0, 3, 8, 15, 25, 60))) %>%
  ggplot(
         aes(x = recall_simu, y = recall_real_cardeorig, 
             fill = sites_per_clone)) +
  geom_abline(slope = 1, intercept = 0, colour = "gray40", linetype = 2) +
  geom_smooth(aes(group = 1), method = "lm", colour = "firebrick") +
  geom_label_repel(aes(label = labs), fill = "gray90", size = 4,
                   nudge_x = 0.05, nudge_y = -0.1) +
  geom_point(size = 3, shape = 21) +
  xlim(0, 1) + ylim(0, 1) +
  scale_fill_manual(name = "mean\n# variants\nper clonal\nbranch", 
                    values = magma(6)[-1]) +
  guides(colour = FALSE, group = FALSE) +
  xlab("Assignment rate: simulated") +
  ylab("Assignment rate: observed") +
  theme_cowplot(font_size = 16)

ggsave("figures/simulations/assign_rate_obs_v_sim.png", plot = p,
       height = 4.5, width = 5)
ggsave("figures/simulations/assign_rate_obs_v_sim.pdf", plot = p,
       height = 4.5, width = 5)

ggsave("figures/simulations/assign_rate_obs_v_sim_wide.png", plot = p,
       height = 4.5, width = 6.5)
ggsave("figures/simulations/assign_rate_obs_v_sim_wide.pdf", plot = p,
       height = 4.5, width = 6.5)
p

ggsave("figures/simulations/assign_rate_obs_v_sim_skinny.png", 
       plot = p + theme(legend.position = "bottom"),
       height = 5.5, width = 5)
ggsave("figures/simulations/assign_rate_obs_v_sim_skinny.pdf", 
       plot = p + theme(legend.position = "bottom"),
       height = 5.5, width = 5)

summary(lm(recall_real_cardeorig ~ recall_simu, data = df))
```

### Cardelino-relax

```{r}
pp <- df %>%
  dplyr::mutate(labs = ifelse(line == "joxm", "joxm", "")) %>%
  dplyr::mutate(sites_per_clone = cut(n_sites / pmax(n_clone - 1, 1), 
                                      breaks = c(0, 3, 8, 15, 25, 60))) %>%
  ggplot(
         aes(x = recall_simu, y = recall_real_carderelax, 
             fill = sites_per_clone)) +
  geom_abline(slope = 1, intercept = 0, colour = "gray40", linetype = 2) +
  geom_smooth(aes(group = 1), method = "lm", colour = "firebrick") +
  geom_label_repel(aes(label = labs), fill = "gray90", size = 4,
                   nudge_x = 0.05, nudge_y = -0.1) +
  geom_point(size = 3, shape = 21) +
  xlim(0, 1) + ylim(0, 1) +
  scale_fill_manual(name = "mean\n# variants\nper clonal\nbranch", 
                    values = magma(6)[-1]) +
  guides(colour = FALSE, group = FALSE) +
  xlab("Assignment rate: simulated") +
  ylab("Assignment rate: observed") +
  theme_cowplot(font_size = 12)

ggsave("figures/simulations/carderelax_assign_rate_obs_v_sim.png", plot = pp,
       height = 4.5, width = 5)
ggsave("figures/simulations/carderelax_assign_rate_obs_v_sim.pdf", plot = pp,
       height = 4.5, width = 5)

ggsave("figures/simulations/carderelax_assign_rate_obs_v_sim_wide.png", plot = pp,
       height = 4.5, width = 6.5)
ggsave("figures/simulations/carderelax_assign_rate_obs_v_sim_wide.pdf", plot = pp,
       height = 4.5, width = 6.5)
pp

ggsave("figures/simulations/carderelax_assign_rate_obs_v_sim_skinny.png", 
       plot = pp + theme(legend.position = "bottom"),
       height = 6.5, width = 6)
ggsave("figures/simulations/carderelax_assign_rate_obs_v_sim_skinny.pdf", 
       plot = pp + theme(legend.position = "bottom"),
       height = 6.5, width = 6)

summary(lm(recall_real_carderelax ~ recall_simu, data = df))
```


## Plot simulation precision-recall curve

```{r}
df %>%
  dplyr::mutate(sites_per_clone = cut(n_sites / n_clone, 
                                      breaks = c(0, 5, 10, 20, 40))) %>%
  ggplot(
         aes(x = recall_simu, y = precision_simu, 
             fill = sites_per_clone)) +
  geom_hline(yintercept = 0.85, colour = "gray40", linetype = 2) +
  geom_smooth(aes(group = 1), method = "lm", colour = "firebrick") +
  geom_point(size = 3, shape = 21) +
  xlim(0, 1) + ylim(0, 1) +
  scale_fill_manual(name = "mean\n# variants\nper clone", 
                    values = magma(5)[-1]) +
  guides(colour = FALSE, group = FALSE) +
  xlab("Assignment rate (recall)") +
  ylab("Precision")


ggsave("figures/simulations/sim_precision_v_recall.png", 
       height = 4.5, width = 5.5)
ggsave("figures/simulations/sim_precision_v_recall.pdf", 
       height = 4.5, width = 5.5)

```

## Clone statistics

Table showing the number of lines with 2, 3 and 4 clones.

```{r}
table(df$n_clone)
```

Summary of the average number of mutations per clonal branch across lines.

```{r}
summary(df$n_sites / (df$n_clone - 1))
```

