---
title: "Tree and probability heatmap plots across all lines"
author: "Davis J. McCarthy"
site: workflowr::wflow_site
---


## Load libraries and data

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.height = 24, fig.width = 24)
library(tidyverse)
library(scater)
library(ggridges)
library(GenomicRanges)
library(RColorBrewer)
library(edgeR)
library(ggrepel)
library(rlang)
library(limma)
library(org.Hs.eg.db)
library(ggforce)
library(cardelino)
library(cowplot)
library(IHW)
library(viridis)
library(ggthemes)
library(superheat)
library(scico)
options(stringsAsFactors = FALSE)
```

Load donors and callset.

```{r}
params <- list()
params$callset <- "filt_lenient.cell_coverage_sites"
fls <- list.files("data/sces")
fls <- fls[grepl(paste0("carderelax.", params$callset), fls)]
lines <- gsub(".*ce_([a-z]+)_with_clone_assignments_carderelax.*", "\\1", fls)
```

Load cell-clone assignment results.

```{r load-cell-assign}
cell_assign_list <- list()
for (don in lines) {
    cell_assign_list[[don]] <- readRDS(file.path("data/cell_assignment", 
        paste0("cardelino_results_carderelax.", don, ".", params$callset, ".rds")))
    cat(paste("reading", don, "\n"))
}
```



## Tree and probability heatmaps: Cardelino updated trees

We can plot the updated clonal trees after running *Cardelino* (using the tree
inferred with *Canopy* as a prior) for each line along with the cell-clone 
assignment results from *cardelino*.


```{r plot-tree-cardelino-updated}
tree_probmat_plot_list <- list()

for (don in lines) {
  ## produce clonal tree plot
  fig_tree <- plot_tree(cell_assign_list[[don]]$tree_updated, orient = "v") + 
    xlab("Clonal tree") +
    cardelino:::heatmap.theme(size = 16) +
    ggtitle(don) +
    theme(axis.text.x = element_blank(), axis.title.y = element_text(size = 20))
  
  prob_to_plot <- cell_assign_list[[don]]$prob_mat[
    colnames(sce_unst_list[[don]])[sce_unst_list[[don]]$well_condition == "unstimulated"], ]
  hc <- hclust(dist(prob_to_plot))
  
  clone_ids <- colnames(prob_to_plot)
  clone_frac <- colMeans(prob_to_plot[matrixStats::rowMaxs(prob_to_plot) > 0.5,, drop = FALSE])
  clone_perc <- paste0(clone_ids, ": ", 
                       round(clone_frac*100, digits = 1), "%")
  
  colnames(prob_to_plot) <- clone_perc
  nba.m <- as_data_frame(prob_to_plot[hc$order,]) %>%
    dplyr::mutate(cell = rownames(prob_to_plot[hc$order,])) %>%
    gather(key = "clone", value = "prob", -cell)
  nba.m <- dplyr::mutate(nba.m, cell = factor(
    cell, levels = rownames(prob_to_plot[hc$order,])))
  fig_assign <- ggplot(nba.m, aes(clone, cell, fill = prob)) + 
    geom_tile(show.legend = TRUE) +
    # scale_fill_gradient(low = "white", high = "firebrick4",
    #                     name = "posterior probability of assignment") +
    scico::scale_fill_scico(palette = "oleron", direction = 1) +
    ylab(paste("Single cells")) + 
    cardelino:::heatmap.theme(size = 16) + #cardelino:::pub.theme() +
    theme(axis.title.y = element_text(size = 20), legend.position = "bottom",
          legend.text = element_text(size = 12), legend.key.size = unit(0.05, "npc"))
  
  p <- plot_grid(fig_tree, fig_assign, nrow = 2, rel_heights = c(0.46, 0.52))
  #print(p)
  
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_probmat.png"), height = 10, width = 7.5)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_probmat.pdf"), height = 10, width = 7.5)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_probmat.svg"), height = 10, width = 7.5)
  # 
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_probmat_wide.png"), height = 9, width = 10)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_probmat_wide.pdf"), height = 9, width = 10)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_probmat_wide.svg"), height = 9, width = 10)
  tree_probmat_plot_list[[don]] <- p
}


p_tree_grid <- plot_grid(plotlist = tree_probmat_plot_list, align = "hv", 
                         nrow = 8)
ggsave(paste0("figures/donor_specific/alldonors_tree_updated_probmat_wide.png"), 
       height = 45, width = 32, limitsize = FALSE) 
ggsave(paste0("figures/donor_specific/alldonors_tree_updated_probmat_wide.pdf"), 
       height = 45, width = 32, limitsize = FALSE) 

p_tree_grid
```


## Tree and probability heatmaps: Canopy trees

We can plot the clonal tree inferred with *Canopy* for each donor along with the
cell-clone assignment results from *cardelino*.


```{r plot-tree-canopy}
tree_canopy_probmat_plot_list <- list()

for (don in lines) {
  ## produce clonal tree plot
  fig_tree <- plot_tree(cell_assign_list[[don]]$full_tree, orient = "v") + 
    xlab("Clonal tree") +
    cardelino:::heatmap.theme(size = 16) +
    ggtitle(don) +
    theme(axis.text.x = element_blank(), axis.title.y = element_text(size = 20))
  
  prob_to_plot <- cell_assign_list[[don]]$prob_mat[
    colnames(sce_unst_list[[don]])[sce_unst_list[[don]]$well_condition == "unstimulated"], ]
  hc <- hclust(dist(prob_to_plot))
  
  clone_ids <- colnames(prob_to_plot)
  clone_frac <- colMeans(prob_to_plot[matrixStats::rowMaxs(prob_to_plot) > 0.5,, drop = FALSE])
  clone_perc <- paste0(clone_ids, ": ", 
                       round(clone_frac*100, digits = 1), "%")
  
  colnames(prob_to_plot) <- clone_perc
  nba.m <- as_data_frame(prob_to_plot[hc$order,]) %>%
    dplyr::mutate(cell = rownames(prob_to_plot[hc$order,])) %>%
    gather(key = "clone", value = "prob", -cell)
  nba.m <- dplyr::mutate(nba.m, cell = factor(
    cell, levels = rownames(prob_to_plot[hc$order,])))
  fig_assign <- ggplot(nba.m, aes(clone, cell, fill = prob)) + 
    geom_tile(show.legend = TRUE) +
    # scale_fill_gradient(low = "white", high = "firebrick4",
    #                     name = "posterior probability of assignment") +
    scico::scale_fill_scico(palette = "oleron", direction = 1) +
    ylab(paste("Single cells")) + 
    cardelino:::heatmap.theme(size = 16) + #cardelino:::pub.theme() +
    theme(axis.title.y = element_text(size = 20), legend.position = "bottom",
          legend.text = element_text(size = 12), legend.key.size = unit(0.05, "npc"))
  
  p <- plot_grid(fig_tree, fig_assign, nrow = 2, rel_heights = c(0.46, 0.52))
  #print(p)
  
  # ggsave(paste0("figures/donor_specific/", don, "_tree_canopy_probmat.png"), height = 10, width = 7.5)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_canopy_probmat.pdf"), height = 10, width = 7.5)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_canopy_probmat.svg"), height = 10, width = 7.5)
  # 
  # ggsave(paste0("figures/donor_specific/", don, "_tree_canopy_probmat_wide.png"), height = 9, width = 10)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_canopy_probmat_wide.pdf"), height = 9, width = 10)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_canopy_probmat_wide.svg"), height = 9, width = 10)
  tree_canopy_probmat_plot_list[[don]] <- p
}


p_tree_grid_canopy <- plot_grid(plotlist = tree_canopy_probmat_plot_list, align = "hv", 
                         nrow = 8)
ggsave(paste0("figures/donor_specific/alldonors_tree_canopy_probmat_wide.png"), 
       height = 50, width = 32, limitsize = FALSE, plot = p_tree_grid_canopy) 
ggsave(paste0("figures/donor_specific/alldonors_tree_canopy_probmat_wide.pdf"), 
       height = 50, width = 32, limitsize = FALSE, plot = p_tree_grid_canopy) 

p_tree_grid_canopy
```


## Cardelino updated tree and Canopy tree comparison plots

```{r tree-tree-plots}
tree_tree_plot_list <- list()

for (don in lines) {
  ## produce clonal tree plot
  fig_tree1 <- plot_tree(cell_assign_list[[don]]$tree_updated, orient = "v") + 
    xlab("Cardelino updated") +
    cardelino:::heatmap.theme(size = 16) +
    theme(axis.text.x = element_blank(), axis.title.y = element_text(size = 20))
  
  fig_tree2 <- plot_tree(cell_assign_list[[don]]$full_tree, orient = "v") + 
    xlab("Canopy tree") +
    cardelino:::heatmap.theme(size = 16) +
    ggtitle(don) +
    theme(axis.text.x = element_blank(), axis.title.y = element_text(size = 20))
  
  p <- plot_grid(fig_tree2, fig_tree1, nrow = 2, rel_heights = c(0.52, 0.46))
  #print(p)
  
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_vs_tree_canopy.png"), height = 10, width = 7.5)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_vs_tree_canopy.pdf"), height = 10, width = 7.5)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_vs_tree_canopy.svg"), height = 10, width = 7.5)
  # 
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_vs_tree_canopy_wide.png"), height = 9, width = 10)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_vs_tree_canopy_wide.pdf"), height = 9, width = 10)
  # ggsave(paste0("figures/donor_specific/", don, "_tree_updated_vs_tree_canopy_wide.svg"), height = 9, width = 10)
  tree_tree_plot_list[[don]] <- p
}

p_tree_tree_grid <- plot_grid(plotlist = tree_tree_plot_list, align = "hv", 
                         nrow = 8)
ggsave(paste0("figures/donor_specific/alldonors_tree_updated_vs_tree_canopy_wide.png"), 
       height = 45, width = 32, limitsize = FALSE, plot = p_tree_tree_grid) 
ggsave(paste0("figures/donor_specific/alldonors_tree_updated_vs_tree_canopy_wide.pdf"), 
       height = 45, width = 32, limitsize = FALSE, plot = p_tree_tree_grid) 

p_tree_tree_grid
```



## Configuration-difference plots 

```{r config-diff-plots}
config_diff_plot_list <- list()

for (don in lines) {
  ## produce clonal tree plot
  
  p <- plot_config_diffs(cell_assign_list[[don]]$Config_prob,
                         cell_assign_list[[don]]$full_tree$Z) +
    ggtitle(don) +
    theme(legend.key.width = unit(0.3, "in")) +
    scale_fill_gradient2(low = "dodgerblue4", mid = "white",
                         high = "firebrick4", midpoint = 0, space = "Lab",
                         na.value = "grey50", guide = "colourbar",
                         aesthetics = "fill", name = "Config diffs",
                         limits = c(-1, 1))
  
  # ggsave(paste0("figures/donor_specific/", don, "_config-diff_cardelino-canopy.png"), height = 8, width = 6)
  # ggsave(paste0("figures/donor_specific/", don, "_config-diff_cardelino-canopy.pdf"), height = 8, width = 6)
  # ggsave(paste0("figures/donor_specific/", don, "_config-diff_cardelino-canopy.svg"), height = 8, width = 6)
  
  config_diff_plot_list[[don]] <- p
}

p_config_diff_grid <- plot_grid(plotlist = config_diff_plot_list, align = "hv",
                         nrow = 4)
ggsave(paste0("figures/donor_specific/alldonors_config-diff_cardelino-canopy.png"),
       height = 20, width = 28, limitsize = FALSE, plot = p_config_diff_grid)
ggsave(paste0("figures/donor_specific/alldonors_config-diff_cardelino-canopy.pdf"),
       height = 20, width = 28, limitsize = FALSE, plot = p_config_diff_grid)

p_config_diff_grid
```






